### 2017-03-15 graphes.tex
save work

\documentclass[french,a4paper]{article}
\usepackage{microtype}
\usepackage{polyglossia}
\usepackage{geometry}
\setmainfont{Tribun ADF Std}
\begin{document}

\section{Graphe ``temporel'' / d'activation}
On définit le graphe dirigé acyclique $G_a=(V_a,E_a)$.
\begin{itemize}
    \item $V_a$ : points d'interaction.
    \item $E_a$ : écoulement entre les points (contraintes temporelles).
\end{itemize}

\section{Graphe hiérarchique}
Graphe de dépendances de données : une fonction a besoin de l'exécution d'une fonction précédente.

Optimisation : passer directement les valeurs. Mais en attendant, on fait du f1 -> store et read -> f2.

\section{Connection entre nœuds}
\begin{itemize}
\item Cas 1 : "optimiste" : on essaye de toujours envoyer les données
\item Cas 2 : "pessimiste" : on n'envoie les données que lorsque tous les nœuds du graphe sont actifs
\item Cas 3 : "retard" : les données sont bufferisées pour envoyer lorsqu'un objet subséquent devient actif. 
Note : que se passe-t-il si on a plusieurs utilisateurs subséquents ?
 Est-ce qu'on reprend du début à chaque fois (le pointeur est dans l'objet situé au bout de l'arête), ou de là ou on s'est arrêté de lire (le pointeur est dans l'objet situé à l'origine de l'arête).
\item Cas 4 : envoyer en parallèle ?
\item Cas 5 : ce qu'il y a actuellement : les nœuds suivants remplacent les résultats des noeuds précédents.
\end{itemize}

Pour pouvoir avoir les différents types de connections dans un seul scénario, il faut avoir moyen de définir des ensembles de processus qui vont pouvoir être liés entre eux.

Dans un groupe on peut définir : 
\begin{itemize}
    \item Une politique.
    \item Des liens qui manifestent la volonté de l'utilisateur.
\end{itemize}

Cas ou on a une variable interne au système : on est toujours soit dans le case 2 ou 3.

Cas ou l'utilisateur doit spécifier explicitement les relations de données : 3, 5

Spécification des sous-graphes

Définir l'interface utilisateur
\end{document}

### 2017-03-15 graphes.tex
minor

@@
 
+\section{Modélisation par des dataflow existants}
+\subsection{Cas du retard}
+On a juste besoin d'ajouter un nœud.
+
+\subsection{Cas 1 : séquence opportuniste}
+Les mettre en série. Idem, est-ce suffisant ?
+
+\subsection{Cas 2 : séquence pessimiste}
+
+\subsection{Cas 4 : Parallèle (utilité?)}
+
+\subsection{Cas 5 : remplacement}
+Les mettre en parallèle, le dernier est celui qu'on veut remplacer (compter que des effets de bord peuvent arriver, donc il doit y avoir un ordre.
+Est-ce suffisant ?
+Ou bien faut-il tout voir dans un seul dataflow...)
+
+** Réfléchir à des examples standard qu'on veut **
+
 Spécification des sous-graphes

### 2017-03-16 article-icmc2017.tex
work

\documentclass[a4paper,twocolumns]{article}
\usepackage{microtype}
\usepackage{polyglossia}
\usepackage{geometry}
\setmainfont{Tribun ADF Std}

\title{Temporal dataflow graphs and their application to generative music}

\begin{document}
	\maketitle

	\section{Introduction}
	\section{Presentation}
	\subsection{Graph definitions}
	\subsubsection{Why multiple graphs ?}
	\subsubsection{Hierarchy}

	\subsection{Inner data and outer data}
	Inner data : reduces the possibilities

	Outer data : have to care about overwriting

	\subsection{Connection types}
	\subsubsection{Opportunistic}
	An execution of a sub-graph can happen whenever any of the nodes of a sub-graph are executing. 
	An order has to be provided.
	\subsubsection{Pessimistic}
	An execution of a sub-graph can only happen when all the nodes of the sub-graph are executing.
	\subsubsection{Delayed}
	A connection between an output and an input is delayed through bufferisation in a queue.

	The bufferisation can behave in two way :
	\begin{itemize}
		\item Readers of the buffer always start from the same point (the beginning of the previous function in the callback chain).
		\item Readers of the buffer continue from the latest read position.
		Question : what happens in case of multiple read ?
	\end{itemize}

	\subsubsection{Parallelized}
	All the sub-graphs are run with the same input environment. 
	The resulting values put in the environment are undefined, but it is guaranteed that all values returned will have been set.

	\subsubsection{Replaced}
	All the sub-graphs are run with the same input environment. 
	The resulting values put in the environment are those of the latest graph execution.

	\section{Meaning}
	Divide the time in abstract segmetn according to the activation graph; gives us a set of equations for each variable.
	\section{Implementation}
	\section{Examples}
\end{document}

### 2017-03-19 article-icmc2017.tex
minor
work

@@
 	\section{Introduction}
-	\section{Presentation}
-	\subsection{Graph definitions}
-	\subsubsection{Why multiple graphs ?}
-	\subsubsection{Hierarchy}
+    This work explores the association of temporal semantics to standard data flow graphs used in music and signal processing.
 
-	\subsection{Inner data and outer data}
-	Inner data : reduces the possibilities
+    We analyze the meanings that can be given to a graphical representation of processes executing in time.
 
-	Outer data : have to care about overwriting
+    For instance, given two processes $p_1, p_2$, what can be said of a program where $p_1$ executes from $t=0$ to $t=5$ seconds, and $p_2$ from $t=3$ to $t=6$ when $p_1$ and $p_2$ operate on the same data.
 
-	\subsection{Connection types}
-	\subsubsection{Opportunistic}
+    The simplest strategy can be to only allow execution when all processes are active and have been given a definite order. 
+    While safe from the point of view of software execution, we believe that other execution policies that leverage information from the temporal layout of processes can provide composers with new creative capabilities.
+
+    \subsection{Related works}
+    Citer Kyma
+    Céu
+    Perl
+
+    \subsection{Examples}
+    Video filtering : 
+    For t=0 to t=5 we change the brightness
+
+    For t=3 to t=6 we change the contrast
+
+    \subsection{Temporal formalism}
+    Rappel i-score / ossia
+
+	\section{Graph definitions}
+	\subsection{Why multiple graphs ?}
+	\subsection{Hierarchy}
+
+	\section{Inner data and outer data}
+	Inner data : reduces the possibilities of connection : everything has to be there (explicit)
+
+	Outer data : have to care about overwriting (implicit)
+    Implicit edges allow for features such as pattern-matching : one node outputs to random, unknown other inputs, or to 
+    the global state.
+    For inputs to handle this, we have to provide them with a "can accept" function.
+
+    \section{Where to put the relationships}
+    \subsection{In the cables}
+    \subsection{In the ports}
+    \subsection{In the nodes}
+    \subsection{In the addresses}
+    No temporal behaviour possible
+    \subsection{Global behaviour}
+    No temporal behaviour possible
+    \subsection{Overall resolution}
+    Unspecified by default.
+
+    Hierarchical : cable, port, node, address, global.
+
+    Default global behaviour : glutton.
+
+	\section{Relationships}
+    \subsection{Strict}
+    An execution of a sub-graph can only happen when all the nodes of the sub-graph are executing.
+
+	\subsection{Glutton}
 	An execution of a sub-graph can happen whenever any of the nodes of a sub-graph are executing. 
 	An order has to be provided.
-	\subsubsection{Pessimistic}
-	An execution of a sub-graph can only happen when all the nodes of the sub-graph are executing.
-	\subsubsection{Delayed}
+
+	\subsection{Delayed}
 	A connection between an output and an input is delayed through bufferisation in a queue.
@@
 
-	\subsubsection{Parallelized}
+	\subsection{Parallelized}
 	All the sub-graphs are run with the same input environment. 
@@
 
-	\subsubsection{Replaced}
+    For audio or video streams however, a meaningful behaviour would be to mix all the outputs together.
+
+	\subsection{Replaced}
 	All the sub-graphs are run with the same input environment. 
@@
 
+    \section{Default behaviours for orders}
+    The temporal graph allows to bypass a bit of cable writing.
+
+    \subsection{Manual}
+    The user explicitely draws a cable.
+
+    \subsection{Hierarchical}
+    We follow the hierarchical organization of the temporal graph.
+
+    \subsection{Temporal glutton}
+    If no order is given for the glutton case, the temporal order in which objects are executed
+    will become the order of chaining in the call graph.
+
 	\section{Meaning}
-	Divide the time in abstract segmetn according to the activation graph; gives us a set of equations for each variable.
+	Divide the time in abstract segment according to the activation graph; gives us a set of equations for each variable.
 	\section{Implementation}
-	\section{Examples}
+    \section{Conclusion}
+    Additional consideration : see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
+
+    Explicit cables : optimization.
 \end{document}

### 2017-03-19 article-icmc2017.tex
work

@@
 
-    The simplest strategy can be to only allow execution when all processes are active and have been given a definite order. 
-    While safe from the point of view of software execution, we believe that other execution policies that leverage information from the temporal layout of processes can provide composers with new creative capabilities.
+    An example would be an audio filter : $p_1$ is a low-pass and $p_2$ a reverb.
+
+    The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution : in this case the processes will only execute from $t=3$ to $t=5$ and behave like function composition.
+    While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of processes can provide composers with new creative capabilities.
 

### 2017-03-20 article-icmc2017.tex
save work

@@
-\documentclass[a4paper,twocolumns]{article}
+\documentclass{article}
+
+\usepackage{icmc2017template}
+\usepackage{times}
+\usepackage{ifpdf}
+\usepackage{soul}
 \usepackage{microtype}
 \usepackage{polyglossia}
-\usepackage{geometry}
-\setmainfont{Tribun ADF Std}
+%\usepackage{geometry}
+
+\def\papertitle{Paper Template for ICMC 2017}
+\def\firstauthor{First Author}
+\def\secondauthor{Second Author}
+\def\thirdauthor{Third Author}
+\def\fourthauthor{Fourth Author}
+\def\fifthauthor{Fifth Author}
+\def\sixthauthor{Sixth Author}
+
+\threeauthors
+{0.5in}
+{\firstauthor} {Affiliation1 \\ %
+    {\tt \href{mailto:author1@smcnetwork.org}{author1@smcnetwork.org}}}
+{\secondauthor} {Affiliation2 \\ %
+    {\tt \href{mailto:author2@smcnetwork.org}{author2@smcnetwork.org}}}
+{\thirdauthor} { Affiliation3 \\ %
+    {\tt \href{mailto:author3@smcnetwork.org}{author3@smcnetwork.org}}}
+
+\usepackage[
+pdftitle={\papertitle},
+pdfauthor={\firstauthor, \secondauthor, \thirdauthor},
+bookmarksnumbered, % use section numbers with bookmarks
+pdfstartview=XYZ % start with zoom=100% instead of full screen; 
+% especially useful if working with a big screen :-)
+]{hyperref}
+%\pdfcompresslevel=9
+
+\usepackage{graphicx}
+\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+
+\usepackage[figure,table]{hypcap}
+
+\hypersetup{
+    colorlinks,%
+    citecolor=black,%
+    filecolor=black,%
+    linkcolor=black,%
+    urlcolor=black
+}
 
@@
 \begin{document}
+
+    \capstartfalse
     \maketitle
+    \capstarttrue
 
@@
     The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution : in this case the processes will only execute from $t=3$ to $t=5$ and behave like function composition.
-    While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of processes can provide composers with new creative capabilities.
+    While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of the processes can provide composers with new creative capabilities.
 
+    The main problem that arises is that composers may specify relationships between processes, but such processes may not always be active. 
+    We start by making the case for an environment associated with the dataflow graph that contains the values of the inputs and outputs.
+    Then, since a graph may not always be fully active, we show that it is also meaningful to have implicit connections between nodes of the graph, that will use the environment.
+    This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
+    We enumerate four different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment : strict, glutton, parallelized, and replaced.
+    Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph.
     \subsection{Related works}
@@
     \subsection{Examples}
-    Video filtering : 
-    For t=0 to t=5 we change the brightness
-
-    For t=3 to t=6 we change the contrast
 
@@
 	\section{Graph definitions}
-	\subsection{Why multiple graphs ?}
-	\subsection{Hierarchy}
+    We rely on two families of graphs for the execution of the software : 
+    \begin{itemize}
+        \item A dataflow graph between processes : processes have input and output ports that can be connected together to enforce data dependencies.
+        \item A temporal graph, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score.
+    \end{itemize}
+    There may be more than a single temporal graph : they can be arranged hierarchically.
 
-	\section{Inner data and outer data}
+    The execution is driven by a scheduler.
+
+    At each tick, the state of each temporal graph is updated with the following informations : 
+    \begin{itemize}
+        \item Which processes are currently running.
+        \item For how long these processes have been running.
+    \end{itemize}
+
+    This is done according to the various events that occur and may trigger conditions and interaction points present in the temporal score.
+
+    Event-driven vs time-driven vs sound-driven ; each node has a "driver" port which is the time or the sound.
+(Do like LV2 : pass empty buffers).
+Pb : multi-rate ...
+voir méthode pour passer de multi-rate à single rate dans PLOP2006 et Signal Flow Graphs and Data Flow Graphs (3.2, 3.3), dans Handbook of Signal Processing (First Edition)
+
+    % Pour thèse : essayer de spécifier l'ensemble des cas possible / configuraitons d'activation de manière statique à partir d'un scénario donné, même si c'est exponentiel..
+
+	%\subsection{Why multiple graphs ?}
+	%\subsection{Hierarchy}
+
+	\section{Implicit and explicit connections}
 	Inner data : reduces the possibilities of connection : everything has to be there (explicit)
@@
     For inputs to handle this, we have to provide them with a "can accept" function.
-
-    \section{Where to put the relationships}
-    \subsection{In the cables}
-    \subsection{In the ports}
-    \subsection{In the nodes}
-    \subsection{In the addresses}
-    No temporal behaviour possible
-    \subsection{Global behaviour}
-    No temporal behaviour possible
-    \subsection{Overall resolution}
-    Unspecified by default.
-
-    Hierarchical : cable, port, node, address, global.
-
-    Default global behaviour : glutton.
 
@@
 
+    \section{Where to put the relationships}
+    \subsection{In the cables}
+    \subsection{In the ports}
+    \subsection{In the nodes}
+    \subsection{In the addresses}
+    No temporal behaviour possible
+    \subsection{Global behaviour}
+    No temporal behaviour possible
+    \subsection{Overall resolution}
+    Unspecified by default.
+
+    Hierarchical : cable, port, node, address, global.
+
+    Default global behaviour : glutton.
+
     \section{Default behaviours for orders}

### 2017-03-21 article-icmc2017.tex
save work

@@
 %\usepackage{geometry}
+\usepackage{tikz}
+\usepackage{listings}
 
@@
 
-	\section{Implicit and explicit connections}
-	Inner data : reduces the possibilities of connection : everything has to be there (explicit)
+    Throughout this paper, we will use the execution trace in fig.~\ref{fig.simple} as running example. 
+    \begin{figure}
+        \centering
+        \begin{tikzpicture}[scale=0.5]
+        \draw[line width=2pt] (0, 0) -- (0, 2) -- (5, 2) -- (5, 0) -- (0, 0);
+        \draw[line width=2pt] (0 + 3, 0 - 2.5) -- (0 + 3, 2 - 2.5) -- (5 + 3, 2 - 2.5) -- (5 + 3, 0 - 2.5) -- (0 + 3, 0 - 2.5);
 
-	Outer data : have to care about overwriting (implicit)
-    Implicit edges allow for features such as pattern-matching : one node outputs to random, unknown other inputs, or to 
-    the global state.
-    For inputs to handle this, we have to provide them with a "can accept" function.
+        \draw[line width=2pt] 
+        (0, -3.75) -- 
+        (0, -4) -- (3, -4) -- 
+        (3, -3.75) --
+        (3, -4) -- (5, -4) -- 
+        (5, -3.75) -- 
+        (5, -4) -- (8, -4) -- 
+        (8, -3.75);
 
+        \node (t1) at (1.45, -3.45) {\large $t_1$};
+        \node (t2) at (4, -3.45) {\large $t_2$};
+        \node (t3) at (6.5, -3.45) {\large $t_3$};
+
+        \node (f1) at (2.5, 1) {\large $f_1$} ;
+        \node (f2) at (5.5, -1.5) {\large $f_2$} ;
+        \end{tikzpicture}
+        \caption{A possible trace of execution for two processes}
+        \label{fig.simple}
+    \end{figure}
+
+	\section{Environment, implicit and explicit connections}
+    In the traditional data flow paradigm, nodes are linked together through connections in their respective input and output ports.
+
+    We extend this with a permanent environment, that contains a mapping of key-value pairs. 
+    Keys are adresses : \lstinline|/foo/bar| and values can be any kind of data : numbers, strings, etc. 
+    The values of the environment may vary independently of the execution of the program : for instance elements of the environment could be changed through the network, graphical or physical controls.
+
+    We then have to provide a mean to use the values of this environment to the dataflow graph.
+    Thus, we extend the inputs and outputs ports with the following expliciteness semantic :
+
+    \begin{itemize}
+        \item An explicit port has been connected by the developer to another port.
+        \item An implicit port has not been connected to another port but may accept any number of adress - value pairs from the environment to perform its work.
+    \end{itemize}
+
+    During a tick, adresses have access to two sets of values : 
+    \begin{itemize}
+        \item The values that were in the environment at the beginning of the tick.
+        \item The values produced by previous nodes in the graph.
+    \end{itemize}
+
+    It is up to the developer of nodes of the graph to choose whether a port can only work when connected explicitely, or with the environment produced from previous nodes, or with the global environment, or both.
+
+    This mechanism allows for new patterns during the creation of a dataflow program : for instance, instead of accepting data from a single source, a port could specify a pattern (see OSC 1.1 paper with XPath) such as : \lstinline|/foo.*/bar/b[a-zA-Z]|.
+    % TODO pattern-matched addresses on the output : seems hard ? how to match it with a "pattern-matched" input port ??
+	%Inner data : reduces the possibilities of connection : everything has to be there (explicit)
+
+	%Outer data : have to care about overwriting (implicit)
+    %Implicit edges allow for features such as pattern-matching : one node outputs to random, unknown other inputs, or to the global state.
+    % For inputs to handle this, we have to provide them with a "can accept" function.
+
+    % TODO sample-precision of the activation / deactivation of a process / node
+    % TODO 
 	\section{Relationships}
+    We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens.
+
     \subsection{Strict}
@@
 
+    That is, if two nodes are compatible, the output of the first node can only go in the input of the second node when both are active.
+
+    For instance, take the case in fig.~\ref{fig.simple} where $f_1$ and $f_2$ both read from $/a$ and write to $/a$.
+
+    Let $t$ be the time of the tick and $e$ the global environment. We assume that $f_1, f_2$ always have access to this information.
+
+    We will get the following set of equations representing the behaviour : 
+    \begin{itemize}
+        \item During $t_0$ : $\emptyset$.
+        \item During $t_1$ : $f_2 \circ f_1 $ or $f_1 \circ f_2$.
+        \item During $t_2$ : $\emptyset$. 
+    \end{itemize}
+
+    The strict execution may happen explicitely or implicitely.
+
 	\subsection{Glutton}
 	An execution of a sub-graph can happen whenever any of the nodes of a sub-graph are executing.
-	An order has to be provided.
+
+    If we take the same case than previously, the behaviour is :
+    \begin{itemize}
+        \item $t_0$ : $f_1$.
+        \item $t_1$ : $f_2 \circ f_1$ or $f_1 \circ f_2$.
+        \item $t_2$ : $f_2$. 
+    \end{itemize}
+
+    The glutton execution may happen explicitely or implicitely.
 
@@
 
-	\subsection{Parallelized}
+    Note : this question has an answer in PLOP2006.
+
+    The delayed execution has be specifieds explicitely for the tokens to be bufferised.
+
+	\subsection{Parallelized and mixed}
 	All the sub-graphs are run with the same input environment. 
-	The resulting values put in the environment are undefined, but it is guaranteed that all values returned will have been set.
+	The resulting values put in the environment are undefined, but it is guaranteed that all values returned will have been set, which may trigger event handlers.
 
@@
 
+    Thus, we have to define summing operations for each of the data types supported.
+
+    We propose for instance : 
+    \begin{itemize}
+        \item For simple parameters such as MIDI control changes or OSC messages : the latest value replaces the previous ones.
+        \item For audio streams : the streams are mixed together.
+        \item For MIDI notes : the notes are played together.
+    \end{itemize}
+
+    The parallelized execution may happen explicitely or implicitely.
+
+    \begin{itemize}
+        \item $t_0$ : $f_1$.
+        \item $t_1$ : $f_1$ and $f_2$ run in parallel, with the same value for $/a$.
+        \item $t_2$ : $f_2$. 
+    \end{itemize}
+
 	\subsection{Replaced}
-	All the sub-graphs are run with the same input environment. 
+	All the nodes are run with the same input environment. 
 	The resulting values put in the environment are those of the latest graph execution.
+
+    \begin{itemize}
+        \item $t_0$ : $f_1$.
+        \item $t_1$ : $f_1$ or $f_2$.
+        \item $t_2$ : $f_2$. 
+    \end{itemize}
+
+    The replaced execution may happen explicitely or implicitely.
 
@@
 
-	\section{Meaning}
+    That is, if the execution happens in the order of the fig.~\ref{fig.simple}, we get the following execution :
+
+    \begin{itemize}
+        \item $t_0$ : $f_1$.
+        \item $t_1$ : $f_2 \circ f_1$.
+        \item $t_2$ : $f_2$. 
+    \end{itemize}
+
+Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get for $t_1$ the execution $f_1 \circ f_2$.
+
+    \begin{figure}
+        \centering
+        \begin{tikzpicture}[scale=0.5]
+        \draw[line width=2pt] (0 + 3, 0) -- (0 + 3, 2) -- (5 + 3, 2) -- (5 + 3, 0) -- (0 + 3, 0);
+        \draw[line width=2pt] (0, 0 - 2.5) -- (0, 2 - 2.5) -- (5, 2 - 2.5) -- (5, 0 - 2.5) -- (0, 0 - 2.5);
+
+        \draw[line width=2pt] 
+        (0, -3.75) -- 
+        (0, -4) -- (3, -4) -- 
+        (3, -3.75) --
+        (3, -4) -- (5, -4) -- 
+        (5, -3.75) -- 
+        (5, -4) -- (8, -4) -- 
+        (8, -3.75);
+
+        \node (t1) at (1.45, -3.45) {\large $t_1$};
+        \node (t2) at (4, -3.45) {\large $t_2$};
+        \node (t3) at (6.5, -3.45) {\large $t_3$};
+
+        \node (f1) at (5.5, 1) {\large $f_1$} ;
+        \node (f2) at (2.5, -1.5) {\large $f_2$} ;
+        \end{tikzpicture}
+        \caption{A possible trace of execution for two processes}
+        \label{fig.simple-reversed}
+    \end{figure}
+
+	\section{Evaluation}
+    We present here the evaluation algorithm that operates at each tick.
+
 	Divide the time in abstract segment according to the activation graph; gives us a set of equations for each variable.

### 2017-03-21 article-icmc2017.tex
save work

@@
 
-    Throughout this paper, we will use the execution trace in fig.~\ref{fig.simple} as running example. 
+    Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example. 
     \begin{figure}
@@
         \label{fig.simple}
+    \end{figure}
+
+    \begin{figure}[h]
+        \centering
+        \begin{tikzpicture}[scale=0.5]
+        \draw[line width=2pt] (0 + 3, 0) -- (0 + 3, 2) -- (5 + 3, 2) -- (5 + 3, 0) -- (0 + 3, 0);
+        \draw[line width=2pt] (0, 0 - 2.5) -- (0, 2 - 2.5) -- (5, 2 - 2.5) -- (5, 0 - 2.5) -- (0, 0 - 2.5);
+
+        \draw[line width=2pt] 
+        (0, -3.75) -- 
+        (0, -4) -- (3, -4) -- 
+        (3, -3.75) --
+        (3, -4) -- (5, -4) -- 
+        (5, -3.75) -- 
+        (5, -4) -- (8, -4) -- 
+        (8, -3.75);
+
+        \node (t1) at (1.45, -3.45) {\large $t_1$};
+        \node (t2) at (4, -3.45) {\large $t_2$};
+        \node (t3) at (6.5, -3.45) {\large $t_3$};
+
+        \node (f1) at (5.5, 1) {\large $f_1$} ;
+        \node (f2) at (2.5, -1.5) {\large $f_2$} ;
+        \end{tikzpicture}
+        \caption{Another trace of execution with the same processes}
+        \label{fig.simple-reversed}
     \end{figure}
@@
 
-    The delayed execution has be specifieds explicitely for the tokens to be bufferised.
+    The delayed execution has be specified explicitely for the tokens to be bufferised.
+
+    If we have a delay connection from $f_1$ to $f_2$, the first call to $f_2$ will use the first value that was produced by $f_1$, and nothing before.
 
@@
     \section{Default behaviours for orders}
-    The temporal graph allows to bypass a bit of cable writing.
+    We study here the various behaviours possible to handle the dependency between nodes of the graph. 
+    We can for instance leverage information provided in the temporal graph to connect the nodes at runtime.
 
-    \subsection{Manual}
-    The user explicitely draws a cable.
+    \subsection{Manual ordering}
+    The user explicitely draws a cable between two ports.
 
     \subsection{Hierarchical}
-    We follow the hierarchical organization of the temporal graph.
+    We follow the hierarchical organization of the temporal graph~: each time constraint has an ordered list of processes, and each time constraint is itself ordered in the scenario.
+This gives an order between processes.
 
-    \subsection{Temporal glutton}
-    If no order is given for the glutton case, the temporal order in which objects are executed
+    \subsection{Temporal}
+    If no order is given, the temporal order in which objects are executed
     will become the order of chaining in the call graph.
@@
 
-    \begin{figure}
-        \centering
-        \begin{tikzpicture}[scale=0.5]
-        \draw[line width=2pt] (0 + 3, 0) -- (0 + 3, 2) -- (5 + 3, 2) -- (5 + 3, 0) -- (0 + 3, 0);
-        \draw[line width=2pt] (0, 0 - 2.5) -- (0, 2 - 2.5) -- (5, 2 - 2.5) -- (5, 0 - 2.5) -- (0, 0 - 2.5);
-
-        \draw[line width=2pt] 
-        (0, -3.75) -- 
-        (0, -4) -- (3, -4) -- 
-        (3, -3.75) --
-        (3, -4) -- (5, -4) -- 
-        (5, -3.75) -- 
-        (5, -4) -- (8, -4) -- 
-        (8, -3.75);
-
-        \node (t1) at (1.45, -3.45) {\large $t_1$};
-        \node (t2) at (4, -3.45) {\large $t_2$};
-        \node (t3) at (6.5, -3.45) {\large $t_3$};
-
-        \node (f1) at (5.5, 1) {\large $f_1$} ;
-        \node (f2) at (2.5, -1.5) {\large $f_2$} ;
-        \end{tikzpicture}
-        \caption{A possible trace of execution for two processes}
-        \label{fig.simple-reversed}
-    \end{figure}
-
-	\section{Evaluation}
+	\section{Evaluation algorithm}
     We present here the evaluation algorithm that operates at each tick.

### 2017-03-22 article-icmc2017.tex
save

@@
 
-    An example would be an audio filter : $p_1$ is a low-pass and $p_2$ a reverb.
+    An example would be an audio filter: $p_1$ is a low-pass and $p_2$ a reverb.
 
-    The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution : in this case the processes will only execute from $t=3$ to $t=5$ and behave like function composition.
+    The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution: in this case the processes will only execute from $t=3$ to $t=5$ and behave like function composition.
     While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of the processes can provide composers with new creative capabilities.
@@
     This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
-    We enumerate four different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment : strict, glutton, parallelized, and replaced.
+    We enumerate four different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment: strict, glutton, parallelized, and replaced.
     Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph.
@@
 	\section{Graph definitions}
-    We rely on two families of graphs for the execution of the software : 
+    We rely on two families of graphs for the execution of the software: 
     \begin{itemize}
-        \item A dataflow graph between processes : processes have input and output ports that can be connected together to enforce data dependencies.
+        \item A dataflow graph between processes: processes have input and output ports that can be connected together to enforce data dependencies.
         \item A temporal graph, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score.
     \end{itemize}
-    There may be more than a single temporal graph : they can be arranged hierarchically.
+    There may be more than a single temporal graph: they can be arranged hierarchically.
 
@@
 
-    At each tick, the state of each temporal graph is updated with the following informations : 
+    At each tick, the state of each temporal graph is updated with the following informations: 
     \begin{itemize}
@@
     Event-driven vs time-driven vs sound-driven ; each node has a "driver" port which is the time or the sound.
-(Do like LV2 : pass empty buffers).
-Pb : multi-rate ...
+(Do like LV2: pass empty buffers).
+Pb: multi-rate ...
 voir méthode pour passer de multi-rate à single rate dans PLOP2006 et Signal Flow Graphs and Data Flow Graphs (3.2, 3.3), dans Handbook of Signal Processing (First Edition)
 
-    % Pour thèse : essayer de spécifier l'ensemble des cas possible / configuraitons d'activation de manière statique à partir d'un scénario donné, même si c'est exponentiel..
+    % Pour thèse: essayer de spécifier l'ensemble des cas possible / configuraitons d'activation de manière statique à partir d'un scénario donné, même si c'est exponentiel..
 
@@
     We extend this with a permanent environment, that contains a mapping of key-value pairs. 
-    Keys are adresses : \lstinline|/foo/bar| and values can be any kind of data : numbers, strings, etc. 
-    The values of the environment may vary independently of the execution of the program : for instance elements of the environment could be changed through the network, graphical or physical controls.
+    Keys are adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
+    The values of the environment may vary independently of the execution of the program: for instance elements of the environment could be changed through the network, graphical or physical controls.
 
     We then have to provide a mean to use the values of this environment to the dataflow graph.
-    Thus, we extend the inputs and outputs ports with the following expliciteness semantic :
+    Thus, we extend the inputs and outputs ports with the following expliciteness semantic:
 
@@
 
-    During a tick, adresses have access to two sets of values : 
+    During a tick, adresses have access to two sets of values: 
     \begin{itemize}
@@
 
-    This mechanism allows for new patterns during the creation of a dataflow program : for instance, instead of accepting data from a single source, a port could specify a pattern (see OSC 1.1 paper with XPath) such as : \lstinline|/foo.*/bar/b[a-zA-Z]|.
-    % TODO pattern-matched addresses on the output : seems hard ? how to match it with a "pattern-matched" input port ??
-	%Inner data : reduces the possibilities of connection : everything has to be there (explicit)
+    This mechanism allows for new patterns during the creation of a dataflow program: for instance, instead of accepting data from a single source, a port could specify a pattern (see OSC 1.1 paper with XPath) such as: \lstinline|/foo.*/bar/b[a-zA-Z]|.
+    % TODO pattern-matched addresses on the output: seems hard ? how to match it with a "pattern-matched" input port ??
+	%Inner data: reduces the possibilities of connection: everything has to be there (explicit)
 
-	%Outer data : have to care about overwriting (implicit)
-    %Implicit edges allow for features such as pattern-matching : one node outputs to random, unknown other inputs, or to the global state.
+	%Outer data: have to care about overwriting (implicit)
+    %Implicit edges allow for features such as pattern-matching: one node outputs to random, unknown other inputs, or to the global state.
     % For inputs to handle this, we have to provide them with a "can accept" function.
@@
 
+     NOTE : if we have cables  we shouldn't have the "parameters" spill to other nodes. 
     \subsection{Strict}
@@
 
-    We will get the following set of equations representing the behaviour : 
+    We will get the following set of equations representing the behaviour: 
     \begin{itemize}
-        \item During $t_0$ : $\emptyset$.
-        \item During $t_1$ : $f_2 \circ f_1 $ or $f_1 \circ f_2$.
-        \item During $t_2$ : $\emptyset$. 
+        \item During $t_0$: $\emptyset$.
+        \item During $t_1$: $f_2 \circ f_1 $ or $f_1 \circ f_2$.
+        \item During $t_2$: $\emptyset$. 
     \end{itemize}
@@
 
-    If we take the same case than previously, the behaviour is :
+    If we take the same case than previously, the behaviour is:
     \begin{itemize}
-        \item $t_0$ : $f_1$.
-        \item $t_1$ : $f_2 \circ f_1$ or $f_1 \circ f_2$.
-        \item $t_2$ : $f_2$. 
+        \item $t_0$: $f_1$.
+        \item $t_1$: $f_2 \circ f_1$ or $f_1 \circ f_2$.
+        \item $t_2$: $f_2$. 
     \end{itemize}
@@
 
-	The bufferisation can behave in two way :
+	The bufferisation can behave in two way:
 	\begin{itemize}
@@
 		\item Readers of the buffer continue from the latest read position.
-		Question : what happens in case of multiple read ?
+		Question: what happens in case of multiple read ?
 	\end{itemize}
 
-    Note : this question has an answer in PLOP2006.
+    Note: this question has an answer in PLOP2006.
 
@@
 
-    We propose for instance : 
+    We propose for instance: 
     \begin{itemize}
-        \item For simple parameters such as MIDI control changes or OSC messages : the latest value replaces the previous ones.
-        \item For audio streams : the streams are mixed together.
-        \item For MIDI notes : the notes are played together.
+        \item For simple parameters such as MIDI control changes or OSC messages: the latest value replaces the previous ones.
+        \item For audio streams: the streams are mixed together.
+        \item For MIDI notes: the notes are played together.
     \end{itemize}
@@
     \begin{itemize}
-        \item $t_0$ : $f_1$.
-        \item $t_1$ : $f_1$ and $f_2$ run in parallel, with the same value for $/a$.
-        \item $t_2$ : $f_2$. 
+        \item $t_0$: $f_1$.
+        \item $t_1$: $f_1$ and $f_2$ run in parallel, with the same value for $/a$.
+        \item $t_2$: $f_2$. 
     \end{itemize}
@@
     \begin{itemize}
-        \item $t_0$ : $f_1$.
-        \item $t_1$ : $f_1$ or $f_2$.
-        \item $t_2$ : $f_2$. 
+        \item $t_0$: $f_1$.
+        \item $t_1$: $f_1$ or $f_2$.
+        \item $t_2$: $f_2$. 
     \end{itemize}
@@
 
-    Hierarchical : cable, port, node, address, global.
+    Hierarchical: cable, port, node, address, global.
 
-    Default global behaviour : glutton.
+    Default global behaviour: glutton.
 
@@
 
-    That is, if the execution happens in the order of the fig.~\ref{fig.simple}, we get the following execution :
+    That is, if the execution happens in the order of the fig.~\ref{fig.simple}, we get the following execution:
 
     \begin{itemize}
-        \item $t_0$ : $f_1$.
-        \item $t_1$ : $f_2 \circ f_1$.
-        \item $t_2$ : $f_2$. 
+        \item $t_0$: $f_1$.
+        \item $t_1$: $f_2 \circ f_1$.
+        \item $t_2$: $f_2$. 
     \end{itemize}
@@
     \section{Conclusion}
-    Additional consideration : see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
+    Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
 
-    Explicit cables : optimization.
+    Explicit cables: optimization.
 \end{document}

### 2017-03-22 article-icmc2017.tex
save

@@
 \usepackage{icmc2017template}
+\usepackage{amssymb}
 \usepackage{times}
@@
 \def\thirdauthor{Third Author}
-\def\fourthauthor{Fourth Author}
-\def\fifthauthor{Fifth Author}
-\def\sixthauthor{Sixth Author}
+
+\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
 
@@
 
-\title{Temporal dataflow graphs and their application to generative music}
+\title{Temporal dataflows and their application to generative music}
 
@@
     \capstarttrue
+    \begin{abstract}
+        contenu...
+    \end{abstract}
 
@@
     Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph.
+
     \subsection{Related works}
@@
     Perl
+    Boolean dataflows (http://pop-art.inrialpes.fr/~fradet/PDFs/RR8333.pdf)
 
@@
 
-	\section{Graph definitions}
-    We rely on two families of graphs for the execution of the software: 
+    \section{Definitions}
+    We call process an entity that consists in a state, an associated function $f$, any number of inputs and output ports, and an activation status.
+    $f$ has a single argument $t \in \mathbb{R^+}$.
+    The execution of $f$ may read any number of tokens from the inputs and write any number of tokens to the outputs.
+
+    We rely on two families of graphs for defining the execution of a program:
+
     \begin{itemize}
-        \item A dataflow graph between processes: processes have input and output ports that can be connected together to enforce data dependencies.
-        \item A temporal graph, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score.
+        \item Synchronous dataflows between processes: input and output ports can be connected together to enforce data dependencies.
+        \item Temporal graphs, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score. 
+        For coherency, we call these graphs timeflows.
     \end{itemize}
-    There may be more than a single temporal graph: they can be arranged hierarchically.
 
-    The execution is driven by a scheduler.
+    A program consists in the sequential and parallel execution of multiple processes, specified by a single dataflow, and at least one timeflow.
 
+    There may be more than a single timeflow: they are defined as processes themselves, which provides them with time information.
+    They have no input or output.
+
+    Each time constraint has an associated set of references to processes.
+    For each process in the dataflow, there is exactly one reference to this process in the set of all timeflows.
+
+    The execution of a program is synchronous and driven by a scheduler.
     At each tick, the state of each temporal graph is updated with the following informations:
+
     \begin{itemize}
-        \item Which processes are currently running.
-        \item For how long these processes have been running.
+        \item Which processes are currently activated.
+        \item For how long these processes have been activated.
     \end{itemize}
@@
 
-    Event-driven vs time-driven vs sound-driven ; each node has a "driver" port which is the time or the sound.
+    Only activated processes may be called in a given tick.
+
+    % In a given tick, every active process will be executed. ?
+    %Event-driven vs time-driven vs sound-driven ; each node has a "driver" port which is the time or the sound.
 (Do like LV2: pass empty buffers).
@@
         \begin{tikzpicture}[scale=0.5]
-        \draw[line width=2pt] (0, 0) -- (0, 2) -- (5, 2) -- (5, 0) -- (0, 0);
-        \draw[line width=2pt] (0 + 3, 0 - 2.5) -- (0 + 3, 2 - 2.5) -- (5 + 3, 2 - 2.5) -- (5 + 3, 0 - 2.5) -- (0 + 3, 0 - 2.5);
+        \draw[line width=2pt] (0, 0) rectangle (5, 2);
+        \draw[line width=2pt] (0 + 3, 0 - 2.5) rectangle (5 + 3, 2 - 2.5);
 
@@
         \begin{tikzpicture}[scale=0.5]
-        \draw[line width=2pt] (0 + 3, 0) -- (0 + 3, 2) -- (5 + 3, 2) -- (5 + 3, 0) -- (0 + 3, 0);
-        \draw[line width=2pt] (0, 0 - 2.5) -- (0, 2 - 2.5) -- (5, 2 - 2.5) -- (5, 0 - 2.5) -- (0, 0 - 2.5);
+
+        \draw[line width=2pt] (0 + 3, 0) rectangle (5 + 3, 2);
+        \draw[line width=2pt] (0, 0 - 2.5) rectangle (5, 2 - 2.5);
 
@@
 
-    We extend this with a permanent environment, that contains a mapping of key-value pairs. 
-    Keys are adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
-    The values of the environment may vary independently of the execution of the program: for instance elements of the environment could be changed through the network, graphical or physical controls.
+    We extend this with a permanent external environment, that contains a mapping of key-value pairs. 
+    Keys are specified with OSC-like\cite{osc} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
+    The values of the environment may vary independently of the execution of the program: for instance elements of the environment could be changed asynchronously through the network, graphical widgets or physical controls.
 

### 2017-03-23 article-icmc2017.tex
save work

@@
     \subsection{Temporal formalism}
-    Rappel i-score / ossia
+
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.45\textwidth]{images/iscore-example.eps}
+        \caption{Part of an i-score scenario, showcasing the temporal syntax used. 
+            A full horizontal line means that the time must not be interrupted, 
+            while a dashed horizontal line means that the time of this time constraint can be interrupted to proceed 
+            to the following parts of the score according to an external event.}
+        \label{fig.iscore-example}
+    \end{figure}
 
@@
     Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example.
+
+    Finally, we will call developers the persons implementing dataflow-based software and users, authors, or composers the persons using dataflow-based software to implement their own music algorithms.
+
     \begin{figure}
@@
 
-        \node (t1) at (1.45, -3.45) {\large $t_1$};
-        \node (t2) at (4, -3.45) {\large $t_2$};
-        \node (t3) at (6.5, -3.45) {\large $t_3$};
+        \node (t1) at (1.45, -3.45) {\large $t_0$};
+        \node (t2) at (4, -3.45) {\large $t_1$};
+        \node (t3) at (6.5, -3.45) {\large $t_2$};
 
@@
 
-    \begin{figure}[h]
+    \begin{figure}
         \centering
@@
     \section{Environment, implicit and explicit connections}
-    In the traditional data flow paradigm, nodes are linked together through connections in their respective input and output ports.
+    In the data flow paradigm, nodes are linked together through connections in their respective input and output ports.
 
-    We extend this with a permanent external environment, that contains a mapping of key-value pairs. 
+    We introduce a permanent external environment, that contains a mapping of key-value pairs.
     Keys are specified with OSC-like\cite{osc} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
-    The values of the environment may vary independently of the execution of the program: for instance elements of the environment could be changed asynchronously through the network, graphical widgets or physical controls.
+    The values of the environment may vary independently of the execution of the program: for instance elements of the environment may be changed asynchronously through the network, graphical widgets or physical controls.
 
-    We then have to provide a mean to use the values of this environment to the dataflow graph.
+    We then have to provide a mean to use the values of this environment to the dataflow.
     Thus, we extend the inputs and outputs ports with the following expliciteness semantic:
@@
     \begin{itemize}
-        \item An explicit port has been connected by the developer to another port.
-        \item An implicit port has not been connected to another port but may accept any number of adress - value pairs from the environment to perform its work.
+        \item An explicit port has been connected manually to another port.
+        An input port may only read tokens from the output port it is connected to.
+
+        \item An implicit port has not been connected to another port. 
+        It may read and write to any number of adresses from the environment to perform its work.
+        Every write done by a node is inserted in a sub-environment scoped to the duration of the tick.
     \end{itemize}
@@
         \item The values that were in the environment at the beginning of the tick.
-        \item The values produced by previous nodes in the graph.
+        \item The values produced by previous implicit output ports in the dataflow.
+        Section~\ref{sec.relationships} presents the various policies that nodes can leverage to 
+        use these values.
     \end{itemize}
 
-    It is up to the developer of nodes of the graph to choose whether a port can only work when connected explicitely, or with the environment produced from previous nodes, or with the global environment, or both.
+    It is up to the implementor of nodes of the graph to choose whether a port can only work: 
+    \begin{itemize}
+        \item When connected explicitely,
+        \item With the environment produced from previous nodes,
+        \item With the global environment,
+        \item With both of the previous.
+    \end{itemize}
 
-    This mechanism allows for new patterns during the creation of a dataflow program: for instance, instead of accepting data from a single source, a port could specify a pattern (see OSC 1.1 paper with XPath) such as: \lstinline|/foo.*/bar/b[a-zA-Z]|.
+    This mechanism allows for new patterns during the creation of a dataflow program.
+    For instance, instead of accepting data from a single source, a port could specify an input pattern (see OSC 1.1 paper with XPath) such as: \lstinline|/foo.*/bar/b[a-zA-Z]|.
     % TODO pattern-matched addresses on the output: seems hard ? how to match it with a "pattern-matched" input port ??
@@
 	\section{Relationships}
-    We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens.
+    \label{sec.relationships}
+    We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens, whether the production is specified implicitely or explicitely.
 
-     NOTE : if we have cables  we shouldn't have the "parameters" spill to other nodes. 
+    We propose five relationships : strict, glutton, delayed, mixed, and replaced. 
+    For some of these relationships, an ordering on the nodes is necessary. 
+    For this, a set of directed edges between nodes of the dataflow is used to provide this order.
+    Methods to set up these edges are discussed in section~\ref{sec.order}.
+
     \subsection{Strict}
@@
 
-    Let $t$ be the time of the tick and $e$ the global environment. We assume that $f_1, f_2$ always have access to this information.
+    Let $t$ be the time of the tick and $e$ the global environment. 
+    We assume that $f_1, f_2$ always have access to this information.
 
@@
         \item During $t_0$: $\emptyset$.
-        \item During $t_1$: $f_2 \circ f_1 $ or $f_1 \circ f_2$.
+        \item During $t_1$: $f_2 \circ f_1 $ or $f_1 \circ f_2$ depending on the node ordering.
         \item During $t_2$: $\emptyset$. 
@@
 
-    The glutton execution may happen explicitely or implicitely.
+    The glutton execution may only happen when nodes are engaged in an implicit relationship.
 
@@
 
-	The bufferisation can behave in two way:
+	The delayed execution can behave in two ways:
 	\begin{itemize}
-		\item Readers of the buffer always start from the same point (the beginning of the previous function in the callback chain).
+		\item Readers of the buffer always start from the same point : the beginning of the previous function in the callback chain.
 		\item Readers of the buffer continue from the latest read position.
-		Question: what happens in case of multiple read ?
+        In this case, the developer has to enforce that a single connection flowing from an output would be active at a given time, to prevent invalid reads .%Question: what happens in case of multiple read ?
 	\end{itemize}
 
-    Note: this question has an answer in PLOP2006.
+    %Note: this question has an answer in PLOP2006.
 
     The delayed execution has be specified explicitely for the tokens to be bufferised.
+    It would be possible
 
-    If we have a delay connection from $f_1$ to $f_2$, the first call to $f_2$ will use the first value that was produced by $f_1$, and nothing before.
+    If we have a delay connection from $f_1$ to $f_2$, the first call to $f_2$ will use the first value that was produced by $f_1$.
+    If the execution order happens to be reversed, such as in the case of fig.~\ref{fig.simple-reversed}, nothing will be read, and the execution will only happen during $t_1$.
 
@@
 
-    We propose for instance: 
+    We propose for instance the following operations: 
     \begin{itemize}
         \item For simple parameters such as MIDI control changes or OSC messages: the latest value replaces the previous ones.
-        \item For audio streams: the streams are mixed together.
+        \item For audio streams: two alternatives : the streams are mixed together, or the streams are put in additional channels.
         \item For MIDI notes: the notes are played together.
@@
 	All the nodes are run with the same input environment. 
-	The resulting values put in the environment are those of the latest graph execution.
+	The resulting values put in the environment are those of the latest node executing in the following dependent nodes.
+
+    Software authors may still have to execute the previous nodes if they happen to produce side-effects.
 
@@
 
-    \section{Where to put the relationships}
+    \section{Specifying the relationships}
+    A remaining problem is the specification of the kind of relationship between two ports, and, more precisely, 
+    where this relationship should be specified.
+
+    We study the implication of bearing such annotations for each object in the dataflow.
+
     \subsection{In the cables}
+    The most precise way to set-up dependency-related information would be in 
+    the cables connecting two ports.
+
+    It is also the only way to set-up a delayed connection.
+
+    However, it would also be the most tedious for users if the patch changes often.
+
     \subsection{In the ports}
@@
     \section{Default behaviours for orders}
+    \label{sec.order}
     We study here the various behaviours possible to handle the dependency between nodes of the graph.
-    We can for instance leverage information provided in the temporal graph to connect the nodes at runtime.
+    information provided provided in the timeflow can be leveraged to connect the nodes at runtime in a fashion 
+    that would make sense to composers \& authors.
+
+    By default, any non-ordered node would strive to be scheduled at the earliest possible time.
+
+    Connections between ports imply an ordering between nodes, hence we only consider the ordering between implicitely connected nodes of the dataflow.
 
     \subsection{Manual ordering}
-    The user explicitely draws a cable between two ports.
+    In this case, the author explicitely specifies an order by setting up an edge between two nodes. 
+    This is the slowest process, but which gives the most precise specification.
 
     \subsection{Hierarchical}
-    We follow the hierarchical organization of the temporal graph~: each time constraint has an ordered list of processes, and each time constraint is itself ordered in the scenario.
-This gives an order between processes.
+    We follow the hierarchical organization of the timeflows~: each time constraint has an ordered list of processes, and each time constraint is itself ordered relative to others in a given timeflow.
+    This gives a natural order between processes.
 
     \subsection{Temporal}
-    If no order is given, the temporal order in which objects are executed
-    will become the order of chaining in the call graph.
+    In this case, the temporal order in which objects are executed
+    will become the order of chaining in the dataflow.
 
-    That is, if the execution happens in the order of the fig.~\ref{fig.simple}, we get the following execution:
+    That is, if a glutton execution happens in the order of the fig.~\ref{fig.simple}, we get the following execution:
 
@@
 
-    Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get for $t_1$ the execution $f_1 \circ f_2$.
+    Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get:
 
-	\section{Evaluation algorithm}
-    We present here the evaluation algorithm that operates at each tick.
+    \begin{itemize}
+        \item $t_0$: $f_2$.
+        \item $t_1$: $f_1 \circ f_2$.
+        \item $t_2$: $f_1$. 
+    \end{itemize}
 
-	Divide the time in abstract segment according to the activation graph; gives us a set of equations for each variable.
+	%\section{Evaluation algorithm}
+    %We present here the evaluation algorithm that operates at each tick.
+
+	%Divide the time in abstract segment according to the activation graph; gives us a set of equations for each variable.
 	\section{Implementation}

### 2017-03-23 article-icmc2017.tex
save

@@
 \usepackage{polyglossia}
-%\usepackage{geometry}
+
 \usepackage{tikz}
 \usepackage{listings}
-
-\def\papertitle{Paper Template for ICMC 2017}
+\usepackage{subcaption}
+\def\papertitle{Temporal dataflows and their application to generative music}
 \def\firstauthor{First Author}
@@
 
-\title{Temporal dataflows and their application to generative music}
+\title{\papertitle}
 
@@
     \section{Introduction}
-    This work explores the association of temporal semantics to standard data flow graphs used in music and signal processing.
+    This work explores the association of macroscopic temporal semantics to standard data flow graphs used in music and signal processing.
 
@@
 
-    An example would be an audio filter: $p_1$ is a low-pass and $p_2$ a reverb.
+    An example would be an audio filter: $p_1$ is a low-pass and $p_2$ a reverb; we want the reverb to activate only after an action of a performer and stop after a few seconds.
 
@@
     This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
-    We enumerate four different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment: strict, glutton, parallelized, and replaced.
-    Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph.
+    We enumerate different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment: strict, glutton, parallelized, and replaced. 
+    Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph, instead of reading the current tokens.
 
@@
 
-    \subsection{Examples}
+%    \subsection{Examples}
 
     \subsection{Temporal formalism}
+    This work implicitely uses the OSSIA formalism to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
+    The elements of its syntax are presented in fig.~\ref{fig.iscore-example}.
 
-    \begin{figure}
+    \begin{figure}[h]
         \centering
@@
 
-    That is, if a glutton execution happens in the order of the fig.~\ref{fig.simple}, we get the following execution:
+    \begin{figure}
+    	\begin{subfigure}{0.15\textwidth }
 
@@
     		\end{itemize}
-
-    Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get:
+    		\caption{Glutton execution in the order of the fig.~\ref{fig.simple}}
+    	\end{subfigure}~
+    \begin{subfigure}{0.15\textwidth}
 
@@
     	\end{itemize}
+    	\caption{
+    		Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get: }
+    \end{subfigure}
+    \caption{}
+    \end{figure}
 

### 2017-03-23 article-icmc2017.tex
save

@@
 
+\setdefaultlanguage{english}
 \usepackage{tikz}
@@
 \usepackage{subcaption}
+\usepackage{booktabs}
+\usepackage{enumitem}\setlist{nosep}
+\setitemize{itemsep=3pt}
+
+\usepackage[backend=biber]{biblatex}
+\addbibresource{biblio.bib}
+
 \def\papertitle{Temporal dataflows and their application to generative music}
@@
 \subsection{Related works}
-    Citer Kyma
-    Céu
-    Perl
-    Boolean dataflows (http://pop-art.inrialpes.fr/~fradet/PDFs/RR8333.pdf)
 
-%    \subsection{Examples}
+OpenMusic reactive\cite{bresson:hal-00965747}
+Timed sequences\cite{garcia:hal-01484077}
+Citer Kyma\cite{scaletti1989kyma}
+Céu\cite{sant2015structured}
+Pearl\cite{halang2001safe}\footnote{Not to be mistaken with the Perl language commonly used for text processing}
+
+Dataflows and especially synchronous dataflows have seen tremendous usage in the music and signal processing community. 
+A list of patterns commonly used when developing dataflow-based music software is presented in~\cite{arumi2006dataflow}.
+Formal semantics for this programming paradigm are given in~\cite{benveniste_data-flow_1993}.
+Specific implementations of dataflow systems are discussed in both editions of the Handbook of Signal Processing Systems\cite{bhattacharyya_handbook_2013}.
+
+Dynamicity in dataflows is generally exprimed on two independent aspects : dynamicity of the data, and of the topology.
+The first relates to the variability on the streams of tokens, while the second is about changes to the structure of the graph. 
+Boolean parametric dataflows\cite{bempelis2015boolean} have been proposed to solve dynamicity of topology, by introducing conditionals at the edges.
 
 \subsection{Temporal formalism}
-    This work implicitely uses the OSSIA formalism to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
+This work implicitely uses the OSSIA formalism~\cite{celerier2015ossia} to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
 The elements of its syntax are presented in fig.~\ref{fig.iscore-example}.
@@
   \centering
-        \includegraphics[width=0.45\textwidth]{images/iscore-example.eps}
+  \includegraphics[width=0.35\textwidth]{images/iscore-example.eps}
   \caption{Part of an i-score scenario, showcasing the temporal syntax used. 
@@
 
+\begin{figure}
+	\centering
+	\begin{tikzpicture}[scale=0.5]
+	 \node[draw, circle] (f1) at (0, 0) { $f_1$ };
+	 \node[draw, circle] (f2) at (3, 0) { $f_2$ }
+	    edge [<-] (f1);
+	\end{tikzpicture}
+	\caption{Another trace of execution with the same processes}
+	\label{fig.simple-reversed}
+\end{figure}
+
 \section{Environment, implicit and explicit connections}
@@
 
-    We introduce a permanent external environment, that contains a mapping of key-value pairs.
-    Keys are specified with OSC-like\cite{osc} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
+This work extends it with a permanent external environment, that contains a mapping of key-value pairs.
+Keys are specified with OSC-like\cite{Freed09featuresand} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
 The values of the environment may vary independently of the execution of the program: for instance elements of the environment may be changed asynchronously through the network, graphical widgets or physical controls.
@@
 
-    It is up to the implementor of nodes of the graph to choose whether a port can only work: 
+Let $\mathrm{ticks}$ be the function that gives the number of ticks in an interval.+
+For a given execution following fig.~\ref{fig.simple}, the evolution of such parameters may look like :
+
+\begin{table}
+	\centering
+	\begin{tabular}{c|ccc}
+		Address & $t_0$ & $t_1$ & $t_2$ \\
+		\midrule
+		Global \lstinline|/a|  & $x$ & $x $ & \\
+		Local \lstinline|/a|   & $\emptyset$ & & \\	
+	\end{tabular}
+	\caption{}
+\end{table}
+
+It is up to the implementor of nodes of the graph to choose whether a port can work: 
 \begin{itemize}
@@
 
-    This mechanism allows for new patterns during the creation of a dataflow program.
-    For instance, instead of accepting data from a single source, a port could specify an input pattern (see OSC 1.1 paper with XPath) such as: \lstinline|/foo.*/bar/b[a-zA-Z]|.
+This mechanism allows for new behaviours during the creation of a dataflow program.
+For instance, instead of accepting data from a single source, a port could specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
 % TODO pattern-matched addresses on the output: seems hard ? how to match it with a "pattern-matched" input port ??
@@
 \subsection{Strict}
-    An execution of a sub-graph can only happen when all the nodes of the sub-graph are executing.
+In a given tick, an execution of a sub-graph can only happen when all the nodes of the sub-graph are activated.
 
@@
 
-    We will get the following set of equations representing the behaviour: 
+We will get the following behaviours during each slice: 
 \begin{itemize}
@@
 
-	\subsection{Delayed}
+\subsection{Delay}
 A connection between an output and an input is delayed through bufferisation in a queue.
@@
 
-	\subsection{Parallelized and mixed}
+\subsection{Parallelization and mixing}
 All the sub-graphs are run with the same input environment. 
@@
 
-	\subsection{Replaced}
+\subsection{Replacement}
 All the nodes are run with the same input environment. 
@@
 \begin{figure}
-    	\begin{subfigure}{0.15\textwidth }
-
+  \begin{subfigure}{0.20\textwidth }    		    		
     \begin{itemize}
@@
   \end{subfigure}~
-    \begin{subfigure}{0.15\textwidth}
+  \begin{subfigure}{0.20\textwidth}
 
@@
     \caption{
-    		Instead, if we are in the case of fig.~\ref{fig.simple-reversed}, we get: }
+    	Glutton execution in the case of fig.~\ref{fig.simple-reversed}}
   \end{subfigure}
@@
 
-	%\section{Evaluation algorithm}
+\section{Evaluation algorithm}
+- Split the tick in steps
+- Each step, go to the next elements (follow the cables) and find all the nodes that can be currently called.
 %We present here the evaluation algorithm that operates at each tick.
@@
 Explicit cables: optimization.
+
+Generalization to not only a single variable of time, but any kind of driving parameters
+\section{Acknowledgements}
+\printbibliography 
 \end{document}

### 2017-03-23 article-icmc2017.tex
work

@@
   \includegraphics[width=0.35\textwidth]{images/iscore-example.eps}
-  \caption{Part of an i-score scenario, showcasing the temporal syntax used. 
+  \caption{Part of an OSSIA scenario, showcasing the temporal syntax used. 
     A full horizontal line means that the time must not be interrupted, 
@@
 
-Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example.
+Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example, as well as the relationships between $f_1$ and $f_2$ present in fig.~\ref{fig.simple-dataflow}.
 
@@
 	\centering
+	\begin{subfigure}{0.16\textwidth}\centering
 		\begin{tikzpicture}[scale=0.5]
-    \draw[line width=2pt] (0, 0) rectangle (5, 2);
-    \draw[line width=2pt] (0 + 3, 0 - 2.5) rectangle (5 + 3, 2 - 2.5);
+		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
+		\node[draw, circle] (f2) at (3, 0) { $f_2$ }
+		edge [<-] (f1);
+		\end{tikzpicture}\caption{}\label{fig.f1tof2}
+	\end{subfigure}~
+	\begin{subfigure}{0.16\textwidth}\centering	
+		\begin{tikzpicture}[scale=0.5]
+		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
+		\node[draw, circle] (f2) at (3, 0) { $f_2$ }
+		edge [->] (f1);
+		\end{tikzpicture}\caption{}
+	\end{subfigure}~	
+	\begin{subfigure}{0.16\textwidth}\centering
+		\begin{tikzpicture}[scale=0.5]
+		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
+		\node[draw, circle] (f2) at (3, 0) { $f_2$ };
+		\end{tikzpicture}\caption{}
+	\end{subfigure}
+	\caption{Possible dataflows between $f_1$ and $f_2$}
+	\label{fig.simple-dataflow}
+\end{figure}
 
-    \draw[line width=2pt] 
+    \begin{figure}
+    	\centering
+\begin{subfigure}{0.25\textwidth}
+  \centering
+  \begin{tikzpicture}[scale=0.25]
+    \draw[] (0, 0) rectangle (5, 2);
+    \draw[] (0 + 3, 0 - 2.5) rectangle (5 + 3, 2 - 2.5);
+
+    \draw[] 
     (0, -3.75) -- 
@@
 
-    \node (t1) at (1.45, -3.45) {\large $t_0$};
-    \node (t2) at (4, -3.45) {\large $t_1$};
-    \node (t3) at (6.5, -3.45) {\large $t_2$};
+    \node (t1) at (1.45, -3.45) {\footnotesize $t_0$};
+    \node (t2) at (4, -3.45) {\footnotesize $t_1$};
+    \node (t3) at (6.5, -3.45) {\footnotesize $t_2$};
 
-    \node (f1) at (2.5, 1) {\large $f_1$} ;
-    \node (f2) at (5.5, -1.5) {\large $f_2$} ;
+    \node (f1) at (2.5, 1) {\footnotesize $f_1$} ;
+    \node (f2) at (5.5, -1.5) {\footnotesize $f_2$} ;
   \end{tikzpicture}
-  \caption{A possible trace of execution for two processes}
+  \caption{A possible trace of execution for $f_1$ and $f_2$}
   \label{fig.simple}
-\end{figure}
+\end{subfigure}~	
+\begin{subfigure}{0.25\textwidth}
+  \centering
+  \begin{tikzpicture}[scale=0.25]
 
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[scale=0.5]
+    \draw[] (0 + 3, 0) rectangle (5 + 3, 2);
+    \draw[] (0, 0 - 2.5) rectangle (5, 2 - 2.5);
 
-    \draw[line width=2pt] (0 + 3, 0) rectangle (5 + 3, 2);
-    \draw[line width=2pt] (0, 0 - 2.5) rectangle (5, 2 - 2.5);
-
-    \draw[line width=2pt] 
+    \draw[] 
     (0, -3.75) -- 
@@
 
-    \node (t1) at (1.45, -3.45) {\large $t_0$};
-    \node (t2) at (4, -3.45) {\large $t_1$};
-    \node (t3) at (6.5, -3.45) {\large $t_2$};
+    \node (t1) at (1.45, -3.45) {\footnotesize $t_0$};
+    \node (t2) at (4, -3.45) {\footnotesize $t_1$};
+    \node (t3) at (6.5, -3.45) {\footnotesize $t_2$};
 
-    \node (f1) at (5.5, 1) {\large $f_1$} ;
-    \node (f2) at (2.5, -1.5) {\large $f_2$} ;
+    \node (f1) at (5.5, 1) {\footnotesize $f_1$} ;
+    \node (f2) at (2.5, -1.5) {\footnotesize $f_2$} ;
   \end{tikzpicture}
-  \caption{Another trace of execution with the same processes}
+  \caption{Another trace of execution}
   \label{fig.simple-reversed}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\begin{tikzpicture}[scale=0.5]
-	 \node[draw, circle] (f1) at (0, 0) { $f_1$ };
-	 \node[draw, circle] (f2) at (3, 0) { $f_2$ }
-	    edge [<-] (f1);
-	\end{tikzpicture}
-	\caption{Another trace of execution with the same processes}
-	\label{fig.simple-reversed}
+\end{subfigure}
+\caption{Process activation traces}
 \end{figure}
@@
 
-During a tick, adresses have access to two sets of values: 
+During a tick, adresses have access to two sets of values, corresponding to scopes: 
 \begin{itemize}
@@
         use these values. 
+        These values are committed to the environment at the end of a tick.
 \end{itemize}
 
-Let $\mathrm{ticks}$ be the function that gives the number of ticks in an interval.+
-For a given execution following fig.~\ref{fig.simple}, the evolution of such parameters may look like :
+For a given execution following fig.~\ref{fig.f1tof2}, the evolution of such parameters would behave as in table~\ref{tbl.evolution}.
 
-\begin{table}
+\begin{table}[h]
 	\centering
-	\begin{tabular}{c|ccc}
-		Address & $t_0$ & $t_1$ & $t_2$ \\
+	\footnotesize
+	\begin{tabular}{c|cccc}
+		Address & Before $f_1$ & After $f_1$ & After $f_2$ & Next tick \\
 		\midrule
-		Global \lstinline|/a|  & $x$ & $x $ & \\
-		Local \lstinline|/a|   & $\emptyset$ & & \\	
+		Global \lstinline|/a|  & $a_0$ & $a_0$ & $a_0$ & $f2(f1(a_0))$\\
+		Local \lstinline|/a|   & $\emptyset$ & $f_1(a_0)$ & $f2(f1(a_0))$ & $\emptyset$ \\	
 	\end{tabular}
-	\caption{}
+	\caption{Evolution of a parameter in the global and local scope}
+	\label{tbl.evolution}
 \end{table}
@@
 
-\subsection{In the cables}
+\subsection{In the cables between ports}
 The most precise way to set-up dependency-related information would be in 
@@
 However, it would also be the most tedious for users if the patch changes often.
+
+\subsection{In the edges between nodes}
 

### 2017-03-23 article-icmc2017.tex
work

@@
 
+Arumi presents in~\cite{arumi2006dataflow} the advantages and drawbacks of both approaches.
+
 %Note: this question has an answer in PLOP2006.
@@
 
-\subsection{In the cables between ports}
+The most obvious candidates are the cables betweens ports, and the dependency edges between nodes of the dataflow.
+
 The most precise way to set-up dependency-related information would be in 
@@
 
-\subsection{In the ports}
-\subsection{In the nodes}
-\subsection{In the addresses}
-No temporal behaviour possible
-\subsection{Global behaviour}
-No temporal behaviour possible
-\subsection{Overall resolution}
-Unspecified by default.
+\subsection{Other}
+Default behaviours can be encoded either globally, per address, or hierarchically in a timeflow.
 
-Hierarchical: cable, port, node, address, global.
+However, temporal behaviours cannot be encoded this way.
 
-Default global behaviour: glutton.
+We choose a default glutton behaviour; however usability studies should be led to determine the most user-friendly default behaviour.
 
@@
 
-\begin{figure}
+With the OSSIA semantics, this order could change at each run of a score.
+
+\begin{figure}[h]
   \begin{subfigure}{0.20\textwidth }    		    		
@@
   \end{subfigure}
-  \caption{}
+  \caption{Executions with a temporal ordering}
 \end{figure}
 
-\section{Evaluation algorithm}
-- Split the tick in steps
-- Each step, go to the next elements (follow the cables) and find all the nodes that can be currently called.
+%\section{Evaluation algorithm}
+%- Split the tick in steps
+%- Each step, go to the next elements (follow the cables) and find all the nodes that can be currently called.
 %We present here the evaluation algorithm that operates at each tick.
@@
 %Divide the time in abstract segment according to the activation graph; gives us a set of equations for each variable.
-\section{Implementation}
+%\section{Implementation}
 \section{Conclusion}
-Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
+%Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
 
-Explicit cables: optimization.
+%Explicit cables: optimization.
 
-Generalization to not only a single variable of time, but any kind of driving parameters
+%Generalization to not only a single variable of time, but any kind of driving parameters
 \section{Acknowledgements}
+Redacted.
 \printbibliography 

### 2017-03-24 article-icmc2017.tex
save

@@
 \usepackage{microtype}
+\usepackage{csquotes}
 \usepackage{polyglossia}
@@
 \usepackage{enumitem}\setlist{nosep}
-\setitemize{itemsep=3pt}
+\setitemize{itemsep=1pt}
 
@@
 
-\def\papertitle{Temporal dataflows and their application to generative music}
+\def\papertitle{Timeflows: extending dataflows in the temporal domain}
 \def\firstauthor{First Author}
@@
 \lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
-
-\threeauthors
-{0.5in}
-{\firstauthor} {Affiliation1 \\ %
-    {\tt \href{mailto:author1@smcnetwork.org}{author1@smcnetwork.org}}}
-{\secondauthor} {Affiliation2 \\ %
-    {\tt \href{mailto:author2@smcnetwork.org}{author2@smcnetwork.org}}}
-{\thirdauthor} { Affiliation3 \\ %
-    {\tt \href{mailto:author3@smcnetwork.org}{author3@smcnetwork.org}}}
 
@@
 \begin{abstract}
-  contenu...
+ % contenu...
 \end{abstract}
@@
 \subsection{Related works}
+There is a long-standing interest in the handling of time in programming languages.
 
-OpenMusic reactive\cite{bresson:hal-00965747}
-Timed sequences\cite{garcia:hal-01484077}
-Citer Kyma\cite{scaletti1989kyma}
-Céu\cite{sant2015structured}
-Pearl\cite{halang2001safe}\footnote{Not to be mistaken with the Perl language commonly used for text processing}
+PEARL90\cite{halang2001safe}\footnote{Not to be mistaken with the Perl language commonly used for text processing} provides temporal primitives allowing for instance to perform loops at a given rate for a given amount of time.
+More recently, Céu has been introduced as a synchronous language with temporal operators, and applications to multimedia\cite{Santos:2016:CLI:2976796.2976856}.
 
-Dataflows and especially synchronous dataflows have seen tremendous usage in the music and signal processing community. 
+OpenMusic is a visual environment which allows to write music by functional composition.
+It has been recently extended with timed sequences allowing to specify evolutions of parameters in time\cite{garcia:hal-01484077}.
+
+Dataflows and especially synchronous dataflows have also seen tremendous usage in the music and signal processing community. 
 A list of patterns commonly used when developing dataflow-based music software is presented in~\cite{arumi2006dataflow}.
-Formal semantics for this programming paradigm are given in~\cite{benveniste_data-flow_1993}.
+Formal semantics are given in~\cite{benveniste_data-flow_1993}.
 Specific implementations of dataflow systems are discussed in both editions of the Handbook of Signal Processing Systems\cite{bhattacharyya_handbook_2013}.
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\begin{subfigure}{0.16\textwidth}\centering
+	\begin{subfigure}{0.15\textwidth}\centering
 		\begin{tikzpicture}[scale=0.5]
@@
 	\end{subfigure}~
-	\begin{subfigure}{0.16\textwidth}\centering	
+	\begin{subfigure}{0.15\textwidth}\centering	
 		\begin{tikzpicture}[scale=0.5]
@@
 	\end{subfigure}~	
-	\begin{subfigure}{0.16\textwidth}\centering
+	\begin{subfigure}{0.15\textwidth}\centering
 		\begin{tikzpicture}[scale=0.5]
@@
 \centering
-\begin{subfigure}{0.25\textwidth}
+\begin{subfigure}{0.22\textwidth}
   \centering
@@
 \end{subfigure}~
-\begin{subfigure}{0.25\textwidth}
+\begin{subfigure}{0.22\textwidth}
   \centering
@@
 \section{Environment, implicit and explicit connections}
-In the data flow paradigm, nodes are linked together through connections in their respective input and output ports.
+In the data flow paradigm generally used in music software, nodes are linked together through connections in their respective input and output ports.
 
@@
         An input port may only read tokens from the output port it is connected to.
-
   \item An implicit port has not been connected to another port. 
@@
         Every write done by a node is inserted in a sub-environment scoped to the duration of the tick.
+        It is up to the developer of the music software to choose whether the requested addresses for each node are specified statically, that is, in the definition of the dataflow, or dynamically.
+        Dynamic behaviour would allow to receive input from a random output at each tick for instance, but makes the program harder to comprehend for authors.
 \end{itemize}
@@
 \begin{itemize}
-  \item The values that were in the environment at the beginning of the tick.
-  \item The values produced by previous implicit output ports in the dataflow.
+  \item Global scope: The values that were in the environment at the beginning of the tick.
+  \item Local scope: The values produced by previous implicit output ports in the dataflow.
         Section~\ref{sec.relationships} presents the various policies that nodes can leverage to 
@@
 	\footnotesize
-	\begin{tabular}{c|cccc}
-		Address & Before $f_1$ & After $f_1$ & After $f_2$ & Next tick \\
+	\begin{tabular}{l|cccc}
+		Address & Tick & After $f_1$ & After $f_2$ & Tick \\
 		\midrule
@@
 	\end{tabular}
-	\caption{Evolution of a parameter in the global and local scope}
+	\caption{Evolution of a parameter in the global and local scope during the execution of a tick. 
+        ``Tick'' indicates the beginning of a new tick; the temporal progression is left-to-right.}
 	\label{tbl.evolution}
@@
 This mechanism allows for new behaviours during the creation of a dataflow program.
-For instance, instead of accepting data from a single source, a port could specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
+For instance, instead of accepting data from a single source, a port can specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
+
+Upon execution, the input stream of the port would contain the list of the matching adresse
+
 % TODO pattern-matched addresses on the output: seems hard ? how to match it with a "pattern-matched" input port ??
@@
 
-%Note: this question has an answer in PLOP2006.
-
 The delayed execution has be specified explicitely for the tokens to be bufferised.
-It would be possible
 
@@
 If the execution order happens to be reversed, such as in the case of fig.~\ref{fig.simple-reversed}, nothing will be read, and the execution will only happen during $t_1$.
+\subsection{Parallelization and mixing}
+The next nodes are run with the same input environment, and summed.
 
-\subsection{Parallelization and mixing}
-All the sub-graphs are run with the same input environment. 
-The resulting values put in the environment are undefined, but it is guaranteed that all values returned will have been set, which may trigger event handlers.
+Summing operations have to be defined for each of the data types supported by the system.
 
-For audio or video streams however, a meaningful behaviour would be to mix all the outputs together.
-
-Thus, we have to define summing operations for each of the data types supported.
-
-We propose for instance the following operations: 
+We propose for instance the following operations for music software: 
 \begin{itemize}
   \item For simple parameters such as MIDI control changes or OSC messages: the latest value replaces the previous ones.
-  \item For audio streams: two alternatives : the streams are mixed together, or the streams are put in additional channels.
+  \item For audio or media streams, the alternatives are to mix the streams together, or to put new streams in additional channels.
   \item For MIDI notes: the notes are played together.
@@
 
-The parallelized execution may happen explicitely or implicitely.
-
 \begin{itemize}
   \item $t_0$: $f_1$.
-  \item $t_1$: $f_1$ and $f_2$ run in parallel, with the same value for $/a$.
+  \item $t_1$: $s(f_1, f_2)$ where $s$ is the summing operation, and $f_1$, $f_2$ are ran with the same input environment
   \item $t_2$: $f_2$. 
 \end{itemize}
+
+The parallelized execution may happen explicitely or implicitely.
 
@@
 
-We study the implication of bearing such annotations for each object in the dataflow.
-
-The most obvious candidates are the cables betweens ports, and the dependency edges between nodes of the dataflow.
-
-The most precise way to set-up dependency-related information would be in 
+The most precise way to set-up dependency-related information would be as an attribute of 
 the cables connecting two ports.
 
-It is also the only way to set-up a delayed connection.
+It is also the only way to set-up a delayed connection, since a buffer may be needed for each connection, but is not possible when an implicit connection is desired.
 
-However, it would also be the most tedious for users if the patch changes often.
+It would also be the most tedious for users if the patch changes often.
 
-\subsection{In the edges between nodes}
+The second possibility is in the dependency edge between nodes; this is compatible with all the relationship kinds, implicit or explicit, at the exception of the delayed connection.
 
-\subsection{Other}
-Default behaviours can be encoded either globally, per address, or hierarchically in a timeflow.
+Finally, default behaviours can be encoded in any hierarchical unit such as the global scope or each time flow could be set-up with a default connection type.
 
-However, temporal behaviours cannot be encoded this way.
-
-We choose a default glutton behaviour; however usability studies should be led to determine the most user-friendly default behaviour.
+We propose a default glutton behaviour; however usability studies should be led to determine the most user-friendly default behaviour.
 
@@
 \label{sec.order}
-We study here the various behaviours possible to handle the dependency between nodes of the graph.
-information provided provided in the timeflow can be leveraged to connect the nodes at runtime in a fashion 
+We assess here the various mechanisms possible to specify the dependencies between nodes of the dataflow.
+Information provided provided in the timeflow can be leveraged to connect the nodes at runtime in a fashion 
 that would make sense to composers \& authors.
@@
 
-\subsection{Hierarchical}
+\subsection{Hierarchical ordering}
 We follow the hierarchical organization of the timeflows~: each time constraint has an ordered list of processes, and each time constraint is itself ordered relative to others in a given timeflow. 
-This gives a natural order between processes.
+This gives a natural order between processes, which does not require the composer to manipulate the dataflow unless a specific data connection has to be established.
 
-\subsection{Temporal}
+\subsection{Temporal ordering}
 In this case, the temporal order in which objects are executed
-will become the order of chaining in the dataflow.
+becomes the order of chaining in the dataflow.
 
-With the OSSIA semantics, this order could change at each run of a score.
+With the OSSIA semantics, this order could change at each run of a score, as showcased in fig.~\ref{fig.temporalordering}.
 
 \begin{figure}[h]
-  \begin{subfigure}{0.20\textwidth }    		    		
+  \begin{subfigure}{0.20\textwidth}
     \begin{itemize}
@@
   \begin{subfigure}{0.20\textwidth}
-
     \begin{itemize}
@@
     \end{itemize}
-    \caption{
-    	Glutton execution in the case of fig.~\ref{fig.simple-reversed}}
+    \caption{Glutton execution in the case of fig.~\ref{fig.simple-reversed}}
   \end{subfigure}
   \caption{Executions with a temporal ordering}
+  \label{fig.temporalordering}
 \end{figure}
@@
 \section{Conclusion}
+Throughought this paper, the effects of adding macro-temporal semantics to dataflows were studied. 
+The first step is to introduce an environment that allows execution of a program to continue if not all the nodes of a dataflow are currently being executed.
+Then, we study the new possibilites offered by the presence of the environment by the different ways to interpret the absence of execution of a node.
+This includes matching nodes together through pattern matching, and distinct policies of execution that can take into account the evolution of the program in time.
+
+An implementation is being developed as an extension to the i-score software. 
+This implementation uses PureData through libpd to provide function execution with multiple ports
+with primitives useful for musical creation.
+
 %Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
 
-%Explicit cables: optimization.
-
 %Generalization to not only a single variable of time, but any kind of driving parameters
-\section{Acknowledgements}
-Redacted.
 \printbibliography 

### 2017-03-27 article-icmc2017.tex
save work

@@
 This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
-We enumerate different behaviours regarding the input and output parameters when the nodes are implicitely connected at runtime using the environment: strict, glutton, parallelized, and replaced. 
-Additionnaly, when the node are explicitely connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph, instead of reading the current tokens.
+We enumerate different behaviors regarding the input and output parameters when the nodes are implicitly connected at runtime using the environment: strict, glutton, parallelized, and replaced. 
+Additionally, when the node are explicitly connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph, instead of reading the current tokens.
 
@@
 
-Dynamicity in dataflows is generally exprimed on two independent aspects : dynamicity of the data, and of the topology.
+Dynamicity in dataflows is generally separated in two independent aspects : dynamicity of the data, and of the topology.
 The first relates to the variability on the streams of tokens, while the second is about changes to the structure of the graph. 
@@
 \subsection{Temporal formalism}
-This work implicitely uses the OSSIA formalism~\cite{celerier2015ossia} to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
+This work implicitly uses the OSSIA formalism~\cite{celerier2015ossia} to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
 The elements of its syntax are presented in fig.~\ref{fig.iscore-example}.
@@
 This work extends it with a permanent external environment, that contains a mapping of key-value pairs.
-Keys are specified with OSC-like\cite{Freed09featuresand} adresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
+Keys are specified with OSC-like\cite{Freed09featuresand} addresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
 The values of the environment may vary independently of the execution of the program: for instance elements of the environment may be changed asynchronously through the network, graphical widgets or physical controls.
@@
 We then have to provide a mean to use the values of this environment to the dataflow.
-Thus, we extend the inputs and outputs ports with the following expliciteness semantic:
+Thus, we extend the inputs and outputs ports with the following explicitness semantic:
 
@@
   \item An implicit port has not been connected to another port. 
-        It may read and write to any number of adresses from the environment to perform its work.
+        It may read and write to any number of addresses from the environment to perform its work.
         Every write done by a node is inserted in a sub-environment scoped to the duration of the tick.
         It is up to the developer of the music software to choose whether the requested addresses for each node are specified statically, that is, in the definition of the dataflow, or dynamically.
-        Dynamic behaviour would allow to receive input from a random output at each tick for instance, but makes the program harder to comprehend for authors.
+        Dynamic behavior would allow to receive input from a random output at each tick for instance, but makes the program harder to comprehend for authors.
 \end{itemize}
 
-During a tick, adresses have access to two sets of values, corresponding to scopes: 
+During a tick, addresses have access to two sets of values, corresponding to scopes: 
 \begin{itemize}
@@
 \begin{itemize}
-  \item When connected explicitely,
+  \item When connected explicitly,
   \item With the environment produced from previous nodes,
@@
 
-This mechanism allows for new behaviours during the creation of a dataflow program.
+This mechanism allows for new behaviors during the creation of a dataflow program.
 For instance, instead of accepting data from a single source, a port can specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
 
-Upon execution, the input stream of the port would contain the list of the matching adresse
+Upon execution, the input stream of the port would contain the list of the matching addresses.
 
@@
 \label{sec.relationships}
-We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens, whether the production is specified implicitely or explicitely.
+We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens, whether the production is specified implicitly or explicitly
 
@@
 
-We will get the following behaviours during each slice: 
+We will get the following behaviors during each slice: 
 \begin{itemize}
@@
 
-The strict execution may happen explicitely or implicitely.
+The strict execution may happen explicitly or implicitly.
 
@@
 
-If we take the same case than previously, the behaviour is:
+If we take the same case than previously, the behavior is:
 \begin{itemize}
@@
 
-The delayed execution has be specified explicitely for the tokens to be bufferised.
+The delayed execution has be specified explicitly for the tokens to be bufferised.
 
@@
 If the execution order happens to be reversed, such as in the case of fig.~\ref{fig.simple-reversed}, nothing will be read, and the execution will only happen during $t_1$.
-\subsection{Parallelization and mixing}
-The next nodes are run with the same input environment, and summed.
+\subsection{Parallelization and reduction}
+The next nodes are run with the same input environment, and summed:
+
+\begin{itemize}
+	\item $t_0$: $f_1$.
+	\item $t_1$: $s(f_1, f_2)$ where $s$ is the reduction operation, and $f_1$, $f_2$ are ran with the same input environment.
+	\item $t_2$: $f_2$. 
+\end{itemize}
 
@@
 
-\begin{itemize}
-  \item $t_0$: $f_1$.
-  \item $t_1$: $s(f_1, f_2)$ where $s$ is the summing operation, and $f_1$, $f_2$ are ran with the same input environment
-  \item $t_2$: $f_2$. 
-\end{itemize}
-
-The parallelized execution may happen explicitely or implicitely.
+The parallelized execution may happen explicitly or implicitly
 
@@
 
-The replaced execution may happen explicitely or implicitely.
+The replaced execution may happen explicitly or implicitly.
 
@@
 
-Finally, default behaviours can be encoded in any hierarchical unit such as the global scope or each time flow could be set-up with a default connection type.
+Finally, default behaviors can be encoded in any hierarchical unit such as the global scope or each time flow could be set-up with a default connection type.
 
-We propose a default glutton behaviour; however usability studies should be led to determine the most user-friendly default behaviour.
+We propose a default glutton behavior; however usability studies should be led to determine the most user-friendly default behavior.
 
-\section{Default behaviours for orders}
+\section{Default behaviors for orders}
 \label{sec.order}
@@
 
-Connections between ports imply an ordering between nodes, hence we only consider the ordering between implicitely connected nodes of the dataflow.
+Connections between ports imply an ordering between nodes, hence we only consider the ordering between implicitly connected nodes of the dataflow.
 
 \subsection{Manual ordering}
-In this case, the author explicitely specifies an order by setting up an edge between two nodes. 
+In this case, the author explicitly specifies an order by setting up an edge between two nodes. 
 This is the slowest process, but which gives the most precise specification.
@@
 \section{Conclusion}
-Throughought this paper, the effects of adding macro-temporal semantics to dataflows were studied. 
+Throughout this paper, the effects of adding macro-temporal semantics to dataflows were studied. 
 The first step is to introduce an environment that allows execution of a program to continue if not all the nodes of a dataflow are currently being executed.
-Then, we study the new possibilites offered by the presence of the environment by the different ways to interpret the absence of execution of a node.
+Then, we study the new possibilities offered by the presence of the environment by the different ways to interpret the absence of execution of a node.
 This includes matching nodes together through pattern matching, and distinct policies of execution that can take into account the evolution of the program in time.
@@
 An implementation is being developed as an extension to the i-score software. 
-This implementation uses PureData through libpd to provide function execution with multiple ports
+This implementation uses PureData via libpd to provide function execution with multiple ports
 with primitives useful for musical creation.
 
+A further goal for this work is to extend the definition not only to work with a single variable of time, but to any kind of activation mechanism: for instance, evolution of a position in a 3D space, etc.
 %Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc

### 2017-03-30 article-icmc2017.tex
save

@@
 \usepackage{polyglossia}
+\usepackage{amsmath}
 
@@
 
-\def\papertitle{Timeflows: extending dataflows in the temporal domain}
+\def\papertitle{Extending dataflows with temporal graphs}
 \def\firstauthor{First Author}
@@
 \def\thirdauthor{Third Author}
+\author{Authors redacted}
+\DeclareMathOperator{\commit}{commit}
+\DeclareMathOperator{\pull}{pull}
 
@@
 \section{Introduction}
-This work explores the association of macroscopic temporal semantics to standard data flow graphs used in music and signal processing.
+This work explores the association of macroscopic temporal semantics with the dataflow graphs generally used in music and signal processing.
 
@@
 This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
-We enumerate different behaviors regarding the input and output parameters when the nodes are implicitly connected at runtime using the environment: strict, glutton, parallelized, and replaced. 
+We enumerate present three different behaviors between the input and output parameters when the nodes are implicitly connected at runtime using the environment: strict, glutton, and delayed. 
 Additionally, when the node are explicitly connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph, instead of reading the current tokens.
@@
 
-Dataflows and especially synchronous dataflows have also seen tremendous usage in the music and signal processing community. 
+Dataflows and especially synchronous dataflows have seen tremendous usage in the music and signal processing community. 
 A list of patterns commonly used when developing dataflow-based music software is presented in~\cite{arumi2006dataflow}.
 Formal semantics are given in~\cite{benveniste_data-flow_1993}.
-Specific implementations of dataflow systems are discussed in both editions of the Handbook of Signal Processing Systems\cite{bhattacharyya_handbook_2013}.
+Specific implementation aspects of dataflow systems are discussed in the Handbook of Signal Processing Systems\cite{bhattacharyya_handbook_2013}.
 
-Dynamicity in dataflows is generally separated in two independent aspects : dynamicity of the data, and of the topology.
+Dynamicity in dataflows is generally separated in two independent aspects: dynamicity of the data, and of the topology.
 The first relates to the variability on the streams of tokens, while the second is about changes to the structure of the graph. 
@@
 \section{Definitions}
-We call process an entity that consists in a state, an associated function $f$, any number of inputs and output ports, and an activation status.
+We call process an entity that consists in a function $f$, an associated state, any number of inputs and output ports, and an activation status.
 $f$ has a single argument $t \in \mathbb{R^+}$.
 The execution of $f$ may read any number of tokens from the inputs and write any number of tokens to the outputs.
+
+The function associated to an inactive process may not be executed.
 
@@
   \item Temporal graphs, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score. 
-  For coherency, we call these graphs timeflows.
+  For coherency, we call these graphs temporal graphs.
 \end{itemize}
 
-A program consists in the sequential and parallel execution of multiple processes, specified by a single dataflow, and at least one timeflow.
+In the following, we consider a program consisting in a dataflow, and a temporal graph.
 
-There may be more than a single timeflow: they are defined as processes themselves, which provides them with time information.
-They have no input or output.
+There may be more than a single temporal graph: they are defined as processes themselves, which provides them with time information.
+A temporal graph does not have inputs or outputs.
 
 Each time constraint has an associated set of references to processes.
-For each process in the dataflow, there is exactly one reference to this process in the set of all timeflows.
+A process may only be executed if it is referenced at least once in a time constraint, 
+at the exception of a special ``main'' process that acts as an entrypoint.
 
@@
 
-This is done according to the various events that occur and may trigger conditions and interaction points present in the temporal score.
+This is done according to the various events that occur and may change the order of activation of processes.
 
-Only activated processes may be called in a given tick.
-
+Processes are then sorted according to their dependencies and activation status, and finally executed.
 % In a given tick, every active process will be executed. ?
@@
 Throughout this paper, we will use the execution traces in fig.~\ref{fig.simple} and fig.~\ref{fig.simple-reversed} as running example, as well as the relationships between $f_1$ and $f_2$ present in fig.~\ref{fig.simple-dataflow}.
-
-Finally, we will call developers the persons implementing dataflow-based software and users, authors, or composers the persons using dataflow-based software to implement their own music algorithms.
 
@@
 	\end{subfigure}
-	\caption{Possible dataflows between $f_1$ and $f_2$}
+	\caption{Possible dataflows between $f_1$ and $f_2$. 
+		As can be seen in fig.~\ref{fig.tgraph}, there are no relationships between $f_1$ and $f_2$ in the temporal graph.}
 	\label{fig.simple-dataflow}
+\end{figure}
+
+\begin{figure}[h]
+
+\centering
+\begin{tikzpicture}[]
+\input{scores/score.tex}
+\end{tikzpicture}
+\caption{The matching temporal graph}
+\label{fig.tgraph}
 \end{figure}
@@
   \begin{tikzpicture}[scale=0.25]
-    \draw[] (0, 0) rectangle (5, 2);
-    \draw[] (0 + 3, 0 - 2.5) rectangle (5 + 3, 2 - 2.5);
+    \draw[] (0, 0) rectangle (5, 2) node[pos=.5] { $f_1$ };
+    \draw[] (0 + 3, 0 - 2.5) rectangle (5 + 3, 2 - 2.5) node[pos=.5] { $f_2$ };
 
@@
     \node (t3) at (6.5, -3.45) {\footnotesize $t_2$};
-
-    \node (f1) at (2.5, 1) {\footnotesize $f_1$} ;
-    \node (f2) at (5.5, -1.5) {\footnotesize $f_2$} ;
   \end{tikzpicture}
@@
 
-    \draw[] (0 + 3, 0) rectangle (5 + 3, 2);
-    \draw[] (0, 0 - 2.5) rectangle (5, 2 - 2.5);
+    \draw[] (0 + 3, 0) rectangle (5 + 3, 2) node[pos=.5] { $f_1$ };
+    \draw[] (0, 0 - 2.5) rectangle (5, 2 - 2.5) node[pos=.5] { $f_2$ };
 
@@
     \node (t3) at (6.5, -3.45) {\footnotesize $t_2$};
-
-    \node (f1) at (5.5, 1) {\footnotesize $f_1$} ;
-    \node (f2) at (2.5, -1.5) {\footnotesize $f_2$} ;
   \end{tikzpicture}
@@
 \end{subfigure}
-\caption{Process activation traces}
+\caption{Process activation traces corresponding to the graph given in \ref{fig.tgraph}}
 \end{figure}
@@
 
-This work extends it with a permanent external environment, that contains a mapping of key-value pairs.
+This work extends the data flow approach with a permanent external environment, that contains a mapping of key-value pairs. 
+Such a pair is called a variable.
+
 Keys are specified with OSC-like\cite{Freed09featuresand} addresses: \lstinline|/foo/bar| and values can be any kind of data: numbers, strings, etc. 
@@
 We then have to provide a mean to use the values of this environment to the dataflow.
-Thus, we extend the inputs and outputs ports with the following explicitness semantic:
+Thus, we extend the input and output ports with the following explicitness semantic:
 
 \begin{itemize}
-  \item An explicit port has been connected manually to another port.
+  \item{An explicit port is a port that has been connected manually to another port.
         An input port may only read tokens from the output port it is connected to.
-  \item An implicit port has not been connected to another port. 
-        It may read and write to any number of addresses from the environment to perform its work.
-        Every write done by a node is inserted in a sub-environment scoped to the duration of the tick.
-        It is up to the developer of the music software to choose whether the requested addresses for each node are specified statically, that is, in the definition of the dataflow, or dynamically.
-        Dynamic behavior would allow to receive input from a random output at each tick for instance, but makes the program harder to comprehend for authors.
+        This connection is static and will not change for the duration of the execution.\footnote{Reactive environments would of course be able to alter such connections at run-time; relevant mechanisms are outside of the scope of this paper.}}
+  \item{An implicit port is a port that has not been connected to another port. 
+        It may dynamically read and write to any number of addresses from the environment to perform its work.}
 \end{itemize}
 
-During a tick, addresses have access to two sets of values, corresponding to scopes: 
+During a tick, ports have access to three sets of values, corresponding to scopes: 
 \begin{itemize}
-  \item Global scope: The values that were in the environment at the beginning of the tick.
-  \item Local scope: The values produced by previous implicit output ports in the dataflow.
-        Section~\ref{sec.relationships} presents the various policies that nodes can leverage to 
-        use these values. 
-        These values are committed to the environment at the end of a tick.
+  \item {Global scope: The values that were in the environment at the beginning of the tick. 
+  These values are accessible to every process.}
+  \item {Local scope: The values produced by previous implicit output ports in the dataflow.
+        Section~\ref{sec.relationships} presents the various policies that nodes can leverage to use these values. 
+        The local scope is cleared at the beginning of a tick. 
+        The values written to it are committed to the global environment at the end of the tick.}
+  \item{Connection scope: the explicit scope between two ports; the data flowing from one out port can only go to input ports it is connected to. This is the explicit case.}
 \end{itemize}
@@
 	\end{tabular}
-	\caption{Evolution of a parameter in the global and local scope during the execution of a tick. 
+	\caption{Evolution of a variable in the global and local scope during the execution of a tick, in the case of an implicit connection. Both $f_1$ and $f_2$ are active. 
         ``Tick'' indicates the beginning of a new tick; the temporal progression is left-to-right.}
@@
 
-It is up to the implementor of nodes of the graph to choose whether a port can work: 
-\begin{itemize}
-  \item When connected explicitly,
-  \item With the environment produced from previous nodes,
-  \item With the global environment,
-  \item With both of the previous.
-\end{itemize}
-
 This mechanism allows for new behaviors during the creation of a dataflow program.
-For instance, instead of accepting data from a single source, a port can specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
-
-Upon execution, the input stream of the port would contain the list of the matching addresses.
 
@@
 \label{sec.relationships}
-We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens, whether the production is specified implicitly or explicitly
+We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens, whether the production is specified implicitly or explicitly, and when taking into account deactivated nodes.
 
-We propose five relationships : strict, glutton, delayed, mixed, and replaced. 
-For some of these relationships, an ordering on the nodes is necessary. 
-For this, a set of directed edges between nodes of the dataflow is used to provide this order.
+We propose three relationships: strict, glutton, and delayed.
+
+These relationships are expressed between ports of two nodes of the dataflow.
+%The last two are expressed between nodes.
+
+Two nodes may not always be explicitly connected to each other through a cable. 
+We have to provide an order between them since execution requires a total ordering of nodes.
+
+For this, an additional set of directed edges between nodes of the dataflow is used to provide this order.
 Methods to set up these edges are discussed in section~\ref{sec.order}.
+\subsection{Strict relationship}
+In a given tick, an execution of a node engaged in a strict relationship with another node depends on the other node being active.
 
-\subsection{Strict}
-In a given tick, an execution of a sub-graph can only happen when all the nodes of the sub-graph are activated.
+For instance, take the case in fig.~\ref{fig.simple} where $f_1$ and $f_2$ both read from $a$ and write to $a$ where $a$ is an address.
+We use the dataflow given in fig.~\ref{fig.f1tof2}.
 
-That is, if two nodes are compatible, the output of the first node can only go in the input of the second node when both are active.
-
-For instance, take the case in fig.~\ref{fig.simple} where $f_1$ and $f_2$ both read from $/a$ and write to $/a$.
-
-Let $t$ be the time of the tick and $e$ the global environment. 
+Let $\commit(a, x)$ be the function that commits a value $x$ to the address $a$ in the local scope, and $\pull(x)$ the function that reads the value of the address $x$ from the global scope.
 We assume that $f_1, f_2$ always have access to this information.
@@
   \item During $t_0$: $\emptyset$.
-  \item During $t_1$: $f_2 \circ f_1 $ or $f_1 \circ f_2$ depending on the node ordering.
+  \item During $t_1$: $\commit(a, (f_2 \circ f_1)(\pull(a)))$.
   \item During $t_2$: $\emptyset$. 
@@
 
-The strict execution may happen explicitly or implicitly.
+A strict relationship between two nodes may only be defined through an explicit connection.
 
-\subsection{Glutton}
-An execution of a sub-graph can happen whenever any of the nodes of a sub-graph are executing.
+This relationship should be used when a computation does not make sense on its own.
+
+%TODO for instance...
+
+\subsection{Glutton relationship}
+An execution of a node will happen even if the nodes it is connected to are not active; instead, data will be read and written from the environment.
 
@@
 \begin{itemize}
-  \item $t_0$: $f_1$.
-  \item $t_1$: $f_2 \circ f_1$ or $f_1 \circ f_2$.
-  \item $t_2$: $f_2$. 
+  \item $t_0$: $\commit(a, f_1(\pull(a)))$.
+  \item $t_1$: $\commit(a, (f_2 \circ f_1)(\pull(a)))$.
+  \item $t_2$: $\commit(a, f_2(\pull(a)))$. 
 \end{itemize}
 
-The glutton execution may only happen when nodes are engaged in an implicit relationship.
+This relationship may be explicit or implicit.
 
-\subsection{Delay}
+In the explicit case, the output of $f_1$ goes to the input of $f_2$ through a cable. 
+If an address has been specified for each port in addition to the explicit connection:
+\begin{itemize}
+	\item If $f_1$ is not active, $f_2$ reads from the local scope instead, or the global scope if the required address is not available.
+	\item If $f_2$ is not active, $f_1$ writes to the local scope instead.
+\end{itemize}
+
+Such a behavior is conceptually similar to a guitarist's pedal board: not all pedals will always be active, but we want the signal to keep flowing even if a pedal is disabled.
+
+\subsection{Delayed relationship}
 A connection between an output and an input is delayed through bufferisation in a queue.
 
-The delayed execution can behave in two ways:
+The delayed connection can behave in two ways:
 \begin{itemize}
-  \item Readers of the buffer always start from the same point : the beginning of the previous function in the callback chain.
+  \item Readers of the buffer always start from the same point: the beginning of the previous function in the callback chain. 
+  The frame pointer would be located in each delayed connection.
   \item Readers of the buffer continue from the latest read position. 
-        In this case, the developer has to enforce that a single connection flowing from an output would be active at a given time, to prevent invalid reads .%Question: what happens in case of multiple read ?
+  The frame pointer would be located in the source port.
+
+  This behavior can be useful when multiple functions should apply successively to a single buffer, as in fig.~\ref{fig.delay2}.
+  However, it would also create concurrent accesses problems if two nodes happened to read an output at the same time.
+
 \end{itemize}
 
-Arumi presents in~\cite{arumi2006dataflow} the advantages and drawbacks of both approaches.
+Arumi presents in~\cite{arumi2006dataflow} the advantages and drawbacks of storing tokens at the output or input ports.
 
-The delayed execution has be specified explicitly for the tokens to be bufferised.
+This relationship has to be specified explicitly to allow for the tokens to be bufferised.
 
 If we have a delay connection from $f_1$ to $f_2$, the first call to $f_2$ will use the first value that was produced by $f_1$.
-If the execution order happens to be reversed, such as in the case of fig.~\ref{fig.simple-reversed}, nothing will be read, and the execution will only happen during $t_1$.
-\subsection{Parallelization and reduction}
-The next nodes are run with the same input environment, and summed:
 
-\begin{itemize}
-	\item $t_0$: $f_1$.
-	\item $t_1$: $s(f_1, f_2)$ where $s$ is the reduction operation, and $f_1$, $f_2$ are ran with the same input environment.
-	\item $t_2$: $f_2$. 
-\end{itemize}
+We define also define a strictness level for the delayed connection. 
+In the strict case, a node will only be able to execute if the source has produced enough tokens.
 
-Summing operations have to be defined for each of the data types supported by the system.
+That is, in fig.~\ref{fig.delay}, $f_2$ would only execute during $t_1$.
 
-We propose for instance the following operations for music software: 
-\begin{itemize}
-  \item For simple parameters such as MIDI control changes or OSC messages: the latest value replaces the previous ones.
-  \item For audio or media streams, the alternatives are to mix the streams together, or to put new streams in additional channels.
-  \item For MIDI notes: the notes are played together.
-\end{itemize}
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[scale=0.45]
+\draw[] (3, 0) rectangle (5, 2) node[pos=.5] { $f_1$ };
+\draw[] (0, 0 - 2.5) rectangle (8, 2 - 2.5) node[pos=.5] { $f_2$ };
 
-The parallelized execution may happen explicitly or implicitly
+\draw[] 
+(0, -3.75) -- 
+(0, -4) -- (3, -4) -- 
+(3, -3.75) --
+(3, -4) -- (5, -4) -- 
+(5, -3.75) -- 
+(5, -4) -- (8, -4) -- 
+(8, -3.75);
 
-\subsection{Replacement}
-All the nodes are run with the same input environment. 
-The resulting values put in the environment are those of the latest node executing in the following dependent nodes.
+\node (t1) at (1.45, -3.45) {\footnotesize $t_0$};
+\node (t2) at (4, -3.45) {\footnotesize $t_1$};
+\node (t3) at (6.5, -3.45) {\footnotesize $t_2$};
+\end{tikzpicture}
+\caption{Another execution trace. $f_2$'s input corresponds to $f_1$'s output, delayed: this creates a causality problem.}
+\label{fig.delay}
+\end{figure}
 
-Software authors may still have to execute the previous nodes if they happen to produce side-effects.
+\begin{figure}[h]
+	\centering
+	\begin{tikzpicture}[scale=0.45]
+	\draw[] (0, 0) rectangle (4, 2) node[pos=.5] { $f_1$ };
+	\draw[] (1, -2.5) rectangle (2, -0.5) node[pos=.5] { $f_2$ };
+	\draw[] (3.5, -2.5) rectangle (4.5, -0.5) node[pos=.5] { $f_3$ };
+	\draw[] (6, -2.5) rectangle (7, -0.5) node[pos=.5] { $f_4$ };
 
-\begin{itemize}
-  \item $t_0$: $f_1$.
-  \item $t_1$: $f_1$ or $f_2$.
-  \item $t_2$: $f_2$. 
-\end{itemize}
+	\end{tikzpicture}
+	\caption{$f_2$, $f_3$ and $f_4$ read from $f_1$ with a delayed connection.}
+	\label{fig.delay2}
+\end{figure}
 
-The replaced execution may happen explicitly or implicitly.
+%If the execution order happens to be reversed, such as in the case of fig.~\ref{fig.simple-reversed}, nothing will be read, and the execution will only happen during $t_1$.
 
-\section{Specifying the relationships}
-A remaining problem is the specification of the kind of relationship between two ports, and, more precisely, 
-where this relationship should be specified.
-
-The most precise way to set-up dependency-related information would be as an attribute of 
-the cables connecting two ports.
-
-It is also the only way to set-up a delayed connection, since a buffer may be needed for each connection, but is not possible when an implicit connection is desired.
-
-It would also be the most tedious for users if the patch changes often.
-
-The second possibility is in the dependency edge between nodes; this is compatible with all the relationship kinds, implicit or explicit, at the exception of the delayed connection.
-
-Finally, default behaviors can be encoded in any hierarchical unit such as the global scope or each time flow could be set-up with a default connection type.
-
-We propose a default glutton behavior; however usability studies should be led to determine the most user-friendly default behavior.
+%\subsection{Parallelization and reduction}
+%In the case where multiple nodes with the same implicit inputs, we 
+%Nodes are parallelized run with the same input environment, and summed:
+%
+%\begin{itemize}
+%	\item $t_0$: $f_1$.
+%	\item $t_1$: $s(f_1, f_2)$ where $s$ is the reduction operation, and $f_1$, $f_2$ are ran with the same input environment.
+%	\item $t_2$: $f_2$. 
+%\end{itemize}
+%
+%Summing operations have to be defined for each of the data types supported by the system.
+%
+%We propose for instance the following operations for music software: 
+%\begin{itemize}
+%  \item For simple parameters such as MIDI control changes or OSC messages: the latest value replaces the previous ones.
+%  \item For audio or media streams, the alternatives are to mix the streams together, or to put new streams in additional channels.
+%  \item For MIDI notes: the notes are played together.
+%\end{itemize}
+%
+%The parallelized execution may happen explicitly or implicitly 
+%
+%\subsection{Replacement}
+%All the nodes are run with the same input environment. 
+%The resulting values put in the environment are those of the latest node executing in the following dependent nodes.
+%    
+%Software authors may still have to execute the previous nodes if they happen to produce side-effects.
+%    
+%\begin{itemize}
+%  \item $t_0$: $f_1$.
+%  \item $t_1$: $f_1$ or $f_2$.
+%  \item $t_2$: $f_2$. 
+%\end{itemize}
+%
+%The replaced execution may happen explicitly or implicitly.
+%    
+%\section{Specifying the relationships}
+%A remaining problem is the specification of the kind of relationship between two ports, and, more precisely, 
+%where this relationship should be specified.
+%
+%The most precise way to set-up dependency-related information would be as an attribute of 
+%the cables connecting two ports. 
+%
+%It is also the only way to set-up a delayed connection, since a buffer may be needed for each connection, but is not possible when an implicit connection is desired.
+%
+%It would also be the most tedious for users if the patch changes often.
+%
+%The second possibility is in the dependency edge between nodes; this is compatible with all the relationship kinds, implicit or explicit, at the exception of the delayed connection.
+%
+%Finally,
 
@@
 We assess here the various mechanisms possible to specify the dependencies between nodes of the dataflow.
-Information provided provided in the timeflow can be leveraged to connect the nodes at runtime in a fashion 
+Information provided provided in the temporal graph can be leveraged to connect the nodes at runtime in a fashion 
 that would make sense to composers \& authors.
 
-By default, any non-ordered node would strive to be scheduled at the earliest possible time.
+By default, any node without preceding nodes would strive to be scheduled at the earliest possible time.
 
-Connections between ports imply an ordering between nodes, hence we only consider the ordering between implicitly connected nodes of the dataflow.
+Connections between ports imply a partial ordering between nodes, hence we only consider the ordering between implicitly connected nodes of the dataflow.
 
@@
 \subsection{Hierarchical ordering}
-We follow the hierarchical organization of the timeflows~: each time constraint has an ordered list of processes, and each time constraint is itself ordered relative to others in a given timeflow. 
-This gives a natural order between processes, which does not require the composer to manipulate the dataflow unless a specific data connection has to be established.
+We follow the hierarchical organization of the temporal graphs: each time constraint has an ordered list of processes, and each time constraint is itself ordered relative to others in a given temporal graph.
+This gives an order between processes, which does not require the composer to manipulate the dataflow unless a specific data connection has to be established.
 
@@
 %\section{Implementation}
+
+\section{Implementation considerations}
+Multiple points discussed in this article can be left to the decision of designers of dataflow systems.
+
+%Finally, we will call developers the persons implementing dataflow-based software and users, authors, or composers the persons using dataflow-based software to implement their own music algorithms.
+\subsection{Compromises in dynamicity}
+A first choice that such a designer ought to make would be the level of dynamicity of the system.
+For instance: would nodes of the dataflow have statically, fixed input and output addresses, or would they be defined and routed dynamically at each tick.
+The first case allows for better performance: more functional dependencies could be checked and set-up statically.
+The second enables new behaviors: for instance, a node could write a value to a random parameter at each tick to create fuzzing effects.
+
+A second choice is the use of scopes and the behavior when a given scope does not have the required value.
+If an input port tries to read from an address \lstinline|/a|, would the implementation allow it to read from the global scope if the value is not available in the local scope?
+It is up to the designer of the software to choose whether a port can work with any permutation of ports: 
+\begin{itemize}
+	\item When connected explicitly,
+	\item With the local scope,
+	\item With the global scope.
+\end{itemize}
+
+\subsection{Pattern matching}
+More dynamic behaviors could also be achieved by allowing to use patterns instead of addresses in the port definitions.
+
+For instance, instead of accepting data from a single source, a port can specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
+
+Upon execution, the input stream of the port would contain the list of values corresponding to the matching addresses.
+
+Given a global environment with the variables: 
+\begin{lstlisting}
+/foo/bar/ba     1.0
+/foo.1/bob/bu   2.0
+/foo.2/bar/be  -1.0
+/foo.42/bar/bo -3.0
+\end{lstlisting}
+
+Such a pattern would yield the tokens $-1.0$ and $-3.0$ at the input of the port matching the pattern.
+
+\subsection{Default behaviors}
+Default behaviors can be encoded in any hierarchical unit; for instance, the global scope. 
+Likewise, each temporal graph could also be set-up with a default connection type.
+
+We propose a default glutton behavior; however usability studies should be made to determine the most user-friendly default.
+
+\subsection{Reference implementation}
+An implementation is being developed as an extension to the i-score software. 
+This implementation uses PureData via libpd to provide function execution with multiple ports
+with primitives useful for musical creation.
+
+\section{Applied example: generative music}
+
 \section{Conclusion}
@@
 
-An implementation is being developed as an extension to the i-score software. 
-This implementation uses PureData via libpd to provide function execution with multiple ports
-with primitives useful for musical creation.
+A further goal for this work is to extend the definition not only to work with a single variable of time, but to any kind of activation mechanism: for instance, evolution of a position in a 3D space, etc.
 
-A further goal for this work is to extend the definition not only to work with a single variable of time, but to any kind of activation mechanism: for instance, evolution of a position in a 3D space, etc.
+Multiple dataflows: can be used as distinct variable scopes. Or use temporal graph as hierarchical scoping mechanism. 
 %Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc

### 2017-03-30 scores/score.tex
save

\definecolor{mygrey}{RGB}{100,100,100}
\draw[line width=1pt] (0, 17.1595)  -- (0, 15.6265) ; % TimeNode.0 
\draw[line width=1pt] (1.64502, 17.1595)  -- (1.64502, 17.1595) ; % drag29wine94 
\draw[line width=1pt] (5, 17.1595)  -- (5, 17.1595) ; % slid38maya79 
\draw[line width=1pt] (1.47186, 15.6265)  -- (1.47186, 15.6265) ; % glut76duff98 
\draw[line width=1pt] (4.85554, 15.6265)  -- (4.85554, 15.6265) ; % ouch30vade32 
\draw[dashed,line width=1pt] (0, 17.1595)  -- (1.64502, 17.1595) ; % gaur3skin98 
\draw[line width=1pt] (1.64502, 17.1595)  -- (5, 17.1595) ; % f1 
\draw[line width=1pt] (1.64502, 17.0595) rectangle (5, 16.0595) node[pos=.5] {$f_1$};
\draw[dashed,line width=1pt] (0, 15.6265)  -- (1.47186, 15.6265) ; % mail74chic13 
\draw[line width=1pt] (1.47186, 15.6265)  -- (4.85554, 15.6265) ; % f2 
\draw[line width=1pt] (1.47186, 15.5265) rectangle (4.85554, 14.5265) node[pos=.5] {$f_2$};

\fill[color=mygrey] (0, 17.1595) circle (0.075) ; % State.0 
\fill[color=mygrey] (1.64502, 17.1595) circle (0.075) ; % State.1 
\fill[color=mygrey] (5, 17.1595) circle (0.075) ; % State.2 
\fill[color=mygrey] (0, 15.6265) circle (0.075) ; % State.3 
\fill[color=mygrey] (1.47186, 15.6265) circle (0.075) ; % State.4 
\fill[color=mygrey] (4.85554, 15.6265) circle (0.075) ; % State.5

### 2017-03-30 article-icmc2017.tex
work

@@
 \section{Applied example: generative music}
+We present here an example that leverages patches built with Pure Data\cite{puckette1996pure}.
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{0.21\textwidth}
+		\includegraphics[scale=0.5]{images/pd-1.png}
+		\caption{$f_1$: generates a random note on each bang}
+	\end{subfigure}
+\begin{subfigure}{0.21\textwidth}
+	\includegraphics[scale=0.5]{images/pd-2.png}
+	\caption{$f_2$: generates another random note on each bang}
+\end{subfigure}
 
+\begin{subfigure}{0.21\textwidth}
+	\includegraphics[scale=0.5]{images/pd-3.png}
+	\caption{$f_3$: synthesizes a note, sets the volume}
+\end{subfigure}
+\begin{subfigure}{0.21\textwidth}
+	\includegraphics[scale=0.5]{images/pd-4.png}
+	\caption{$f_4$: applies an high-pass filter}
+\end{subfigure}
+\begin{subfigure}{0.21\textwidth}
+\includegraphics[scale=0.5]{images/pd-5.png}
+\caption{$f_5$: applies a low-pass filter}
+\end{subfigure}
+\caption{The functions we use in our example, as defined in PureData.}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=0.7]{images/nodes.eps}
+
+	\caption{The relationships between instances of $f_{1,..5}$ in the program}
+\end{figure}
+
+\begin{figure}
+	\begin{tikzpicture}
+	\input{scores/score2.tex}
+	\end{tikzpicture}
+	\caption{}
+\end{figure}
 \section{Conclusion}

### 2017-03-31 article-icmc2017.tex
save work

@@
 \usepackage{icmc2017template}
+\usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{times}
-\usepackage{ifpdf}
 \usepackage{soul}
@@
 \usepackage{polyglossia}
-\usepackage{amsmath}
+\setmainfont{Times New Roman}
 
@@
 \DeclareMathOperator{\pull}{pull}
+\usepackage[explicit]{titlesec}
+\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
 
-\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
+\titleformat{name = \section, numberless}[runin]
+{\bfseries}
+{}
+{1em}
+{#1}
 
@@
 \begin{abstract}
- % contenu...
+    Numeric musical and artistic creation is in great part based on well-known patcher software: Max/MSP, PureData, OpenMusic, vvvvv\dots
+
+    Yet, handling temporal evolutions and structures on large scales remains a central problem: for instance, writing a song with introduction, chorus, verse, chorus, parts in Max presents many hurdles to the composer.
+    Hence, multiple software have introduced tracks and automations to provide a temporal control of a given set of parameters in a single or even multiple concurrent time-lines.
+
+    We present an extension to the dataflow paradigm used by patchers that allows to handle the execution of any parts of a computation in time. 
+    That is, instead of sending control messages to existing patchs, they are run in a time-line which orchestrates the data bindings between different parts of the dataflow. 
+    This enables dynamic behaviours when creating temporal compositions. An example is given with Pure Data patches and an i-score scenario.
 \end{abstract}
@@
 
-We analyze the meanings that can be given to a graphical representation of processes executing in time.
+We analyze the meanings that can be given to processes executing in time, when these processes have a data dependency and do not execute at exactly the same time.
 
@@
 The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution: in this case the processes will only execute from $t=3$ to $t=5$ and behave like function composition.
-While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of the processes can provide composers with new creative capabilities.
+While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of the processes can provide composers with new creative capabilities, through an example that uses Pure Data.
 
-The main problem that arises is that composers may specify relationships between processes, but such processes may not always be active. 
-We start by making the case for an environment associated with the dataflow graph that contains the values of the inputs and outputs.
+An environment associated with the dataflow graph is introduced; it contains the values of the inputs and outputs of the processes in this graph.
 Then, since a graph may not always be fully active, we show that it is also meaningful to have implicit connections between nodes of the graph, that will use the environment.
 This allows dynamic routing of the processes according to their temporal order during execution, and extends routing to pattern-matched elements of the environment instead of single values.
-We enumerate present three different behaviors between the input and output parameters when the nodes are implicitly connected at runtime using the environment: strict, glutton, and delayed. 
-Additionally, when the node are explicitly connected, we introduce a delay mechanism that allows a computation to happen on data previously produced by another node of the graph, instead of reading the current tokens.
+Three different behaviors between the input and output parameters of the dataflow nodes are presented: strict, glutton, and delayed.
 
@@
 \subsection{Temporal formalism}
-This work implicitly uses the OSSIA formalism~\cite{celerier2015ossia} to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines). 
+This work implicitly uses the OSSIA formalism~\cite{celerier2015ossia} in its examples to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines).
 The elements of its syntax are presented in fig.~\ref{fig.iscore-example}.
+
+It is important to note that any other system of temporal relationships between processes could be used instead.
 
@@
 
-For instance, instead of accepting data from a single source, a port can specify an input pattern such as: \lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
+For instance, instead of accepting data from a single source, a port can specify an input pattern such as:~\\\lstinline|/foo.*/bar/b[a-zA-Z]|, as proposed in the OSC specification.
 
@@
 \subsection{Reference implementation}
-An implementation is being developed as an extension to the i-score software. 
-This implementation uses PureData via libpd to provide function execution with multiple ports
+An implementation is being developed as an extension to the i-score software\footnote{\url{http://www.i-score.org}}. 
+This implementation uses PureData~\cite{puckette1996pure} through libpd~\cite{brinkmann2011embedding} to provide function execution with multiple ports
 with primitives useful for musical creation.
@@
 \section{Applied example: generative music}
-We present here an example that leverages patches built with Pure Data\cite{puckette1996pure}.
 \begin{figure}[h]
     \centering
-	\begin{subfigure}{0.21\textwidth}
+    \begin{subfigure}{0.21\textwidth}\centering
         \includegraphics[scale=0.5]{images/pd-1.png}
-		\caption{$f_1$: generates a random note on each bang}
+        \caption{$f_1, f'_1$: generates a random note on each bang}
     \end{subfigure}
-\begin{subfigure}{0.21\textwidth}
+    \begin{subfigure}{0.21\textwidth}\centering
         \includegraphics[scale=0.5]{images/pd-2.png}
@@
 
-\begin{subfigure}{0.21\textwidth}
+    \begin{subfigure}{0.21\textwidth}\centering
         \includegraphics[scale=0.5]{images/pd-3.png}
@@
     \end{subfigure}
-\begin{subfigure}{0.21\textwidth}
+    \begin{subfigure}{0.21\textwidth}\centering
         \includegraphics[scale=0.5]{images/pd-4.png}
@@
     \end{subfigure}
-\begin{subfigure}{0.21\textwidth}
+    \begin{subfigure}{0.21\textwidth}\centering
         \includegraphics[scale=0.5]{images/pd-5.png}
-\caption{$f_5$: applies a low-pass filter}
+        \caption{$f_5, f'_5$: applies a low-pass filter}
     \end{subfigure}
     \caption{The functions we use in our example, as defined in PureData.}
+    \label{fig.pdpatches}
 \end{figure}
@@
 
-	\caption{The relationships between instances of $f_{1,..5}$ in the program}
+    \caption{The data relationships between instances of $f_{1,..5}$ in the program.}
+    \label{fig.datapatch}
 \end{figure}
@@
 \begin{figure}
+    \centering
     \begin{tikzpicture}
@@
     \end{tikzpicture}
-	\caption{}
+    \caption{The temporal relationships between instances of $f_{1,..5}$ in the program.}
+    \label{fig.timepatch}
 \end{figure}
+
+We present here an example that leverages patches built with Pure Data\cite{puckette1996pure}.
+Figure~\ref{fig.pdpatches} present the Pd patches used.
+Fig.~\ref{fig.datapatch} presents the data relationships between the Pd patches, with the mapping of inputs and outputs to addresses of the environment.
+The inlets and outlets are ordered so that audio ports come first and message ports come afterwards.
+Finally, fig.~\ref{fig.timepatch} presents the temporal score that orchestrates the execution of the Pd patches.
+
+The execution happens as follows: 
+\begin{enumerate}
+    \item For a few milliseconds, nothing happens.
+    \item $f_3$ is started. However, there is a strict dependency between $f_3$ and $f_4$: no sound is produced; $f_3$ is disabled.
+    \item $f_1$ is started. Since $f_3$ is disabled, it writes \lstinline|/note| to the environment.
+    \item $f_1$ stops, and a few milliseconds afterwards, $f_4$ starts, which allows  $f_3$ to execute. $f_3$ uses the last value of \lstinline|/note| that was put in the global scope by $f_1$.
+    $f_3$ fetches the value of \lstinline|/vol| in the environment and $f_4$ fetches the value of \lstinline|/filt| there as well.
+    Notice that $f_3$'s parent time constraint is dashed: it will execute until it is manually stopped.
+
+    The sound that can be heard is a single tone with some amount of high-pass filtering.
+    \item $f_2$ starts: random notes will be generated and put at the \lstinline|/note| input of $f_3$ directly, bypassing the environment.
+
+    \item $f_5$ starts: the low-pass filter adds itself after the high-pass filter.
+    \item $f_2$ stops: the sound reverts to a continuous tone (the last that was written, again).
+    \item $f'_1$ starts: tones are generated again, but with a faster rhythm.
+    \item $f_4$ stops: there is no sound anymore due to the strict dependency. $f_5$ applies its low-pass filter to silence.
+    \item $f_5$ stops, and just afterwards $f'5$ starts.
+Since $f'5$ was in a delayed relationship with $f_4$, the sound produced corresponds to the sound that was heard at the beginning of $f_4$, with both $f_4$'s high pass filtering, and $f'5$'s low-pass filtering.
+\end{enumerate}
 \section{Conclusion}
@@
 
+Such global environments have been implicitly present in multiple dataflow-based software; however they are not given an explicit existence.
+For instance, Max/MSP externals generally handle their global state through \lstinline|static| C variables.
+Making it explicit allows for better understanding of the execution of a program, and may enable further optimization opportunities.
+
+After discussing the various choices left to the implementor, an example leveraging multiple PureData patches scheduled through an i-score score is explained.
+
 A further goal for this work is to extend the definition not only to work with a single variable of time, but to any kind of activation mechanism: for instance, evolution of a position in a 3D space, etc.
 
-Multiple dataflows: can be used as distinct variable scopes. Or use temporal graph as hierarchical scoping mechanism. 
+The paper mostly centered around a single dataflow.
+A meaning has been proposed for multiple dataflows: they could be used as distinct variable scopes.
+If the temporal graph has hierarchical features, its use as a scoping mechanism should also be envisioned.
+
+\printbibliography
+
 %Additional consideration: see PLOP 2006 article (multi-stream, etc), sample accuracy, etc
@@
 %Generalization to not only a single variable of time, but any kind of driving parameters
-\printbibliography 
+
 \end{document}

### 2017-03-31 article-icmc2017.tex
save work

@@
 \begin{abstract}
-    Numeric musical and artistic creation is in great part based on well-known patcher software: Max/MSP, PureData, OpenMusic, vvvvv\dots
+    Numeric musical and artistic creation is in great part based on well-known patcher softwares: Max/MSP, PureData, OpenMusic, vvvv\dots
 

### 2017-03-31 article-icmc2017.tex
update

@@
 \usepackage{polyglossia}
+\usepackage{footmisc}
 \setmainfont{Times New Roman}
@@
 
-For instance, given two processes $p_1, p_2$, what can be said of a program where $p_1$ executes from $t=0$ to $t=5$ seconds, and $p_2$ from $t=3$ to $t=6$ when $p_1$ and $p_2$ operate on the same data.
+For instance, given two processes $p_1, p_2$, what can be said of a program where $p_1$ executes from $t=0$ to $t=5$ seconds, and $p_2$ from $t=3$ to $t=6$ when $p_1$ and $p_2$ operate on the same data?
 
@@
 
-The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution: in this case the processes will only execute from $t=3$ to $t=5$ and behave like function composition.
+The simplest strategy can be to only allow execution when all processes are active and have been given a definite order of execution: in this case the overall program will only execute during the intersection of the activation of $p_1$ and $p_2$, which goes from $t=3$ to $t=5$.
 While safe from the point of view of software execution, we will show that other execution policies leveraging information from the temporal layout of the processes can provide composers with new creative capabilities, through an example that uses Pure Data.
@@
 Three different behaviors between the input and output parameters of the dataflow nodes are presented: strict, glutton, and delayed.
+
+\begin{figure} [h]
+	\centering
+	\includegraphics[width=0.35\textwidth]{images/iscore-example.eps}
+	\caption{Part of an OSSIA scenario, showcasing the temporal syntax used. 
+		A full horizontal line means that the time must not be interrupted, 
+		while a dashed horizontal line means that the time of this time constraint can be interrupted to proceed 
+		to the following parts of the score according to an external event.}
+	\label{fig.iscore-example}
+\end{figure}
 
@@
 \subsection{Temporal formalism}
+
 This work implicitly uses the OSSIA formalism~\cite{celerier2015ossia} in its examples to provide primitives relative to the evolution of time: time constraints (horizontal lines), time nodes (vertical lines).
@@
 
+\section{Definitions}
 \begin{figure}[h]
 	\centering
-  \includegraphics[width=0.35\textwidth]{images/iscore-example.eps}
-  \caption{Part of an OSSIA scenario, showcasing the temporal syntax used. 
-    A full horizontal line means that the time must not be interrupted, 
-    while a dashed horizontal line means that the time of this time constraint can be interrupted to proceed 
-  to the following parts of the score according to an external event.}
-  \label{fig.iscore-example}
+	\begin{subfigure}{0.15\textwidth}\centering
+		\begin{tikzpicture}[scale=0.5]
+		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
+		\node[draw, circle] (f2) at (3, 0) { $f_2$ }
+		edge [<-] (f1);
+		\end{tikzpicture}\caption{}\label{fig.f1tof2}
+	\end{subfigure}~
+	\begin{subfigure}{0.15\textwidth}\centering	
+		\begin{tikzpicture}[scale=0.5]
+		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
+		\node[draw, circle] (f2) at (3, 0) { $f_2$ }
+		edge [->] (f1);
+		\end{tikzpicture}\caption{}
+	\end{subfigure}~	
+	\begin{subfigure}{0.15\textwidth}\centering
+		\begin{tikzpicture}[scale=0.5]
+		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
+		\node[draw, circle] (f2) at (3, 0) { $f_2$ };
+		\end{tikzpicture}\caption{}
+	\end{subfigure}
+	\caption{Possible dataflows between $f_1$ and $f_2$. 
+		As can be seen in fig.~\ref{fig.tgraph}, there are no relationships between $f_1$ and $f_2$ in the temporal graph.}
+	\label{fig.simple-dataflow}
 \end{figure}
 
-\section{Definitions}
 We call process an entity that consists in a function $f$, an associated state, any number of inputs and output ports, and an activation status.
@@
 \begin{itemize}
-  \item Synchronous dataflows between processes: input and output ports can be connected together to enforce data dependencies.
+  \item Synchronous dataflow graphs between processes:~\\$G_d(V_d, E_d)$ where $V_d$ is the set of processes and $E_d$ the set of data bindings between processes.
+
   \item Temporal graphs, based on the OSSIA formalism: $G_t(V_t, E_t)$ where $V_t$ is the set of time nodes and $E_t$ is the set of time constraints in a score. 
@@
 
-There may be more than a single temporal graph: they are defined as processes themselves, which provides them with time information.
-A temporal graph does not have inputs or outputs.
+A program may consist in more than a single temporal graph: they are defined as processes themselves, which provides them with time information.
 
@@
 \begin{figure}[h]
-	\centering
-	\begin{subfigure}{0.15\textwidth}\centering
-		\begin{tikzpicture}[scale=0.5]
-		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
-		\node[draw, circle] (f2) at (3, 0) { $f_2$ }
-		edge [<-] (f1);
-		\end{tikzpicture}\caption{}\label{fig.f1tof2}
-	\end{subfigure}~
-	\begin{subfigure}{0.15\textwidth}\centering	
-		\begin{tikzpicture}[scale=0.5]
-		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
-		\node[draw, circle] (f2) at (3, 0) { $f_2$ }
-		edge [->] (f1);
-		\end{tikzpicture}\caption{}
-	\end{subfigure}~	
-	\begin{subfigure}{0.15\textwidth}\centering
-		\begin{tikzpicture}[scale=0.5]
-		\node[draw, circle] (f1) at (0, 0) { $f_1$ };
-		\node[draw, circle] (f2) at (3, 0) { $f_2$ };
-		\end{tikzpicture}\caption{}
-	\end{subfigure}
-	\caption{Possible dataflows between $f_1$ and $f_2$. 
-		As can be seen in fig.~\ref{fig.tgraph}, there are no relationships between $f_1$ and $f_2$ in the temporal graph.}
-	\label{fig.simple-dataflow}
-\end{figure}
-
-\begin{figure}[h]
 
@@
 \end{tikzpicture}
-\caption{The matching temporal graph}
+\caption{The example temporal graph we use: both $f_1$ and $f_2$ can start at any time. $f_1$ and $f_2$ both have a given, finite duration.}
 \label{fig.tgraph}
@@
 
-\begin{figure}
+\begin{figure}[htp]
 \centering
@@
 We are interested in the relationships between nodes of the dataflow graph when they produce compatible tokens, whether the production is specified implicitly or explicitly, and when taking into account deactivated nodes.
+That is, given a node of the dataflow executing and producing tokens, we must define which following nodes, if any, will receive the tokens and when will they execute.
 
@@
   \item Readers of the buffer always start from the same point: the beginning of the previous function in the callback chain. 
-  The frame pointer would be located in each delayed connection.
+  The frame pointer would be located in each delayed connection, and would not be shared between processes.
   \item Readers of the buffer continue from the latest read position. 
-  The frame pointer would be located in the source port.
+  The frame pointer would be located in the source port, and would be shared across all processes reading from it.
 
@@
 
-This relationship has to be specified explicitly to allow for the tokens to be bufferised.
+This relationship has to be specified explicitly to allow for the tokens to be buffered.
 
@@
 
-We define also define a strictness level for the delayed connection. 
+We also define a strictness level for the delayed connection. 
 In the strict case, a node will only be able to execute if the source has produced enough tokens.
@@
 \end{tikzpicture}
-\caption{Another execution trace. $f_2$'s input corresponds to $f_1$'s output, delayed: this creates a causality problem.}
+\caption{Another execution trace, stemming from a different score than the one presented in fig.~\ref{fig.tgraph}.
+$f_2$'s input corresponds to $f_1$'s output, delayed: this creates a causality problem.
+}
 \label{fig.delay}
@@
 We assess here the various mechanisms possible to specify the dependencies between nodes of the dataflow.
-Information provided provided in the temporal graph can be leveraged to connect the nodes at runtime in a fashion 
+Information provided in the temporal graph can be leveraged to connect the nodes at runtime in a fashion 
 that would make sense to composers \& authors.
@@
 
-Given a global environment with the variables: 
+Given a global environment with the variables:~\\
+\begin{tabular}{c}
 \begin{lstlisting}
@@
 \end{lstlisting}
-
+\end{tabular}~\\~\\
 Such a pattern would yield the tokens $-1.0$ and $-3.0$ at the input of the port matching the pattern.
@@
         \includegraphics[scale=0.5]{images/pd-1.png}
-        \caption{$f_1, f'_1$: generates a random note on each bang}
+        \caption{$f_1, f'_1$: note generator.}
     \end{subfigure}
@@
         \includegraphics[scale=0.5]{images/pd-2.png}
-        \caption{$f_2$: generates another random note on each bang}
+        \caption{$f_2$: note generator, at a lower tempo.}
     \end{subfigure}
@@
         \includegraphics[scale=0.5]{images/pd-3.png}
-        \caption{$f_3$: synthesizes a note, sets the volume}
+        \caption{$f_3$: synthesizes the input note, sets the volume.}
     \end{subfigure}
+
     \begin{subfigure}{0.21\textwidth}\centering
         \includegraphics[scale=0.5]{images/pd-4.png}
-        \caption{$f_4$: applies an high-pass filter}
+        \caption{$f_4$: high-pass filter.}
     \end{subfigure}
@@
         \includegraphics[scale=0.5]{images/pd-5.png}
-        \caption{$f_5, f'_5$: applies a low-pass filter}
+        \caption{$f_5, f'_5$: low-pass filter.}
     \end{subfigure}
-    \caption{The functions we use in our example, as defined in PureData.}
+    \caption{The functions we use in our example, as defined in Pure Data.}
     \label{fig.pdpatches}
@@
 
-    \caption{The data relationships between instances of $f_{1,..5}$ in the program.}
+    \caption{The dataflow relationships between instances of $f_{1,..5}$ in the program.
+The global environment contains variables for \lstinline|/vol| and \lstinline|/filt| when the execution starts.}
     \label{fig.datapatch}
@@
     The sound that can be heard is a single tone with some amount of high-pass filtering.
-    \item $f_2$ starts: random notes will be generated and put at the \lstinline|/note| input of $f_3$ directly, bypassing the environment.
+    \item $f_2$ starts: random notes will be generated at a lower tempo and put at the \lstinline|/note| input of $f_3$ directly, bypassing the environment.
 
@@
     \item $f_2$ stops: the sound reverts to a continuous tone (the last that was written, again).
-    \item $f'_1$ starts: tones are generated again, but with a faster rhythm.
+    \item $f'_1$ starts: tones are generated again, but with a faster tempo.
     \item $f_4$ stops: there is no sound anymore due to the strict dependency. $f_5$ applies its low-pass filter to silence.

### 2017-07-21 article-icmc2017.tex
add some todos
minor
wip

@@
 \def\papertitle{Extending dataflows with temporal graphs}
-\def\firstauthor{First Author}
-\def\secondauthor{Second Author}
-\def\thirdauthor{Third Author}
-\author{Authors redacted}
+\def\firstauthor{Jean-Michaël Celerier}
+\def\secondauthor{Myriam Desainte-Catherine}
+\def\thirdauthor{Jean-Michel Couturier}
+
+\threeauthors
+{0.5in}
+{\firstauthor} {LaBRI, Blue Yeti \\
+    Univ. Bordeaux, LaBRI, UMR 5800, ~\\ F-33400 Talence, France. \\  Blue Yeti, F-17110 France. \\ jcelerie@labri.fr}
+{\secondauthor} {LaBRI, CNRS \\ Univ.
+Bordeaux, LaBRI, UMR 5800, ~\\ F-33400 Talence, France.
+\\ CNRS, LaBRI, UMR 5800, ~\\F-33400 Talence, France.
+\\  INRIA, F-33400 Talence, France.
+\\ myriam@labri.fr}
+{\thirdauthor} {Blue Yeti, F-17110 France. \\ jmc@blueyeti.fr}
+
 \DeclareMathOperator{\commit}{commit}
@@
 
+%%%%% TODO pour review
+%% Ok: Parler de Bach et du truc du mec de l'IRCAM (dynamic patching en lisp)
+%% Ok: Expliquer que ce n'est pas qu'un mécanisme pour patcher Pd; à venir: objets de i-score.
+%% Présenter ça comme la fusion du modèle flot de données et du modèle séquenceur comme dans Ableton + Max For Live
+%% Différence avec patching dynamique: l'agencement temporel des objets entre eux peut être statique.
+%% Expliquer les raisons: 
+%% Mettre accent sur aspects temporels
+%% Ok:  Expliquer qu'on présente un paradigme, qu'il est implémenté à bas niveau dans la libossia et à haut niveau dans i-score (dont l'intégration avec Pd)
+%% Bof: Accentuer le fait qu'il n'y ait pas besoin de coder
+%% Bof: Définir la notion de dynamicité utilisée.
+%% Bof: "Any other system of temporal relationships" : donner examples. Par ex. Antescofo, etc ?
+%% Ok: Expliquer implémentation: mettre capture d'écran ?
+
+%% User evaluation... bof
 \begin{document}
@@
 \subsection{Related works}
-There is a long-standing interest in the handling of time in programming languages.
+There is a long-standing interest in the handling of time in programming languages, which is intrinsically linked to how the language handles dynamicity.
 
@@
 It has been recently extended with timed sequences allowing to specify evolutions of parameters in time\cite{garcia:hal-01484077}.
+
+Likewise, the Bach library for Max~\cite{agostini2015max} allows to define temporal variations of parameters during the playing of a note by with the mechanism of slots.
+The processes controlled by such parameters are then available to use in the Max patch.
+
+The Max for Live extension to Ableton Live allows to embed Max patches in the Ableton Live sequencer. 
+Through the API provided, one can control the execution of various elements of the sequencer in Max; automations in Live can also be used to send data to Max patches at a given time.
+
+A method for dynamic patching of Max abstractions based on CommmonLisp has been proposed by Thomas Hummel\cite{hummel1994common} to reduce resource usage by enabling and disabling sub-patches at different points in the execution of a program.
+This has the advantage of saving computing power for the active elements of the score.
 
@@
 
-\subsection{Reference implementation}
-An implementation is being developed as an extension to the i-score software\footnote{\url{http://www.i-score.org}}. 
-This implementation uses PureData~\cite{puckette1996pure} through libpd~\cite{brinkmann2011embedding} to provide function execution with multiple ports
-with primitives useful for musical creation.
-
-\section{Applied example: generative music}
-\begin{figure}[h]
+\begin{figure}
     \centering
@@
 
-\begin{figure}[h]
+\subsection{Reference implementation}
+An implementation is being developed as an extension to the OSSIA library\footnote{\url{https://github.com/OSSIA/libossia}} and i-score software\footnote{\url{http://www.i-score.org}}. 
+This implementation uses PureData~\cite{puckette1996pure} through libpd~\cite{brinkmann2011embedding} to provide function execution with multiple ports with primitives useful for musical creation. 
+As such, it allows to work with audio, midi, and messages.
+
+The C++ implementation of the execution algorithm itself, provided in the OSSIA library is not tied in any way to PureData or any other environment: PureData is used as an example of implementation of a node to leverage an easy and well-known method to produce generators and filters with different kinds of inputs and outputs, which allows to cover the whole set of features.
+Other nodes can be implemented through inheritance of a set of C++ classes.
+As such, the processes currently available in i-score (automations, Javascript code, MIDI piano roll) and in i-score-audio (sound file reading, Faust and LV2 effects) are being ported to this new execution model.
+
+In terms of user interface, PureData patches appear as nodes with their inputs and outputs in i-score.
+There are two views: the main view, temporal and hierarchical, which allows to position the nodes in time, draw automations, etc., and the patching view which allows to connect nodes together and set the relationship types.
+The temporal view instead has easily accessible text fields that allow setting input and output addresses for each port.
+
+\section{Applied example: generative music}
+\begin{figure}
     \centering
@@
     \label{fig.datapatch}
-\end{figure}
-
+\end{figure}~
 \begin{figure}
@@
 Such global environments have been implicitly present in multiple dataflow-based software; however they are not given an explicit existence.
-For instance, Max/MSP externals generally handle their global state through \lstinline|static| C variables.
+For instance, Max externals generally handle their global state through \lstinline|static| C variables.
 Making it explicit allows for better understanding of the execution of a program, and may enable further optimization opportunities.
@@
 If the temporal graph has hierarchical features, its use as a scoping mechanism should also be envisioned.
-
+% TODO video / shader
 \printbibliography

### 2017-07-21 article-icmc2017.tex
wip

@@
 \label{fig.delay}
-\end{figure}
-
+\end{figure}\vspace{-5mm}
 \begin{figure}[h]
@@
 \end{figure}
+\vspace{-5mm}
 
@@
 
-The C++ implementation of the execution algorithm itself, provided in the OSSIA library is not tied in any way to PureData or any other environment: PureData is used as an example of implementation of a node to leverage an easy and well-known method to produce generators and filters with different kinds of inputs and outputs, which allows to cover the whole set of features.
+The C++ implementation of the execution algorithm, provided in the OSSIA library, is not tied in any way to PureData or any other environment: PureData is used as an example of implementation of a node to leverage an easy and well-known method to produce generators and filters with different kinds of inputs and outputs, which allows to cover the whole set of features.
 Other nodes can be implemented through inheritance of a set of C++ classes.
@@
     \caption{The dataflow relationships between instances of $f_{1,..5}$ in the program.
-The global environment contains variables for \lstinline|/vol| and \lstinline|/filt| when the execution starts.}
+The global environment contains variables for \lstinline|/vol| and \lstinline|/filt| when the execution starts.
+A PureData patch with similar semantics would use additionally \lstinline|throw~, catch~, send, receive, delwrite~, delread~,tabwrite~, tabread~| between nodes, and UDP \& OSC receivers and senders to read and write from the environment.}
     \label{fig.datapatch}

### 2017-10-13 article-icmc2017.tex
minor

@@
 \begin{abstract}
-    Numeric musical and artistic creation is in great part based on well-known patcher softwares: Max/MSP, PureData, OpenMusic, vvvv\dots
+    Numeric musical and artistic creation is in great part based on well-known patcher softwares: Max, PureData, OpenMusic, vvvv\dots
 
@@
 
-A method for dynamic patching of Max abstractions based on CommmonLisp has been proposed by Thomas Hummel\cite{hummel1994common} to reduce resource usage by enabling and disabling sub-patches at different points in the execution of a program.
+A method for dynamic patching of Max abstractions based on CommonLisp has been proposed by Thomas Hummel\cite{hummel1994common} to reduce resource usage by enabling and disabling sub-patches at different points in the execution of a program.
 This has the advantage of saving computing power for the active elements of the score.
@@
         The values written to it are committed to the global environment at the end of the tick.}
-  \item{Connection scope: the explicit scope between two ports; the data flowing from one out port can only go to input ports it is connected to. This is the explicit case.}
+  \item{Connection scope: the explicit scope between two ports; the data flowing from one output port can only go to input ports it is connected to. This is the explicit case.}
 \end{itemize}
